---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Sample/Study Overview & Geodistnaces
Basic descriptive data for samples and study location. 
```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = T, results = 'hide',
                      fig.align = 'center', fig.width = 8, fig.height = 8)
knitr::opts_knit$set(root.dir = here::here())
```
```{r}
library(tidyverse)
library(sf)
source("R/basics.R")
```

## Sample Characteristics
Samples are combination of adults, children with a full DHS line, and children that were "accidentally" sampled (i.e. not a full DHS line -- occurred when child was slightly older than 5 yrs but a DBS was still taken). 
```{r}
#..............................................................
# import for plotting
#..............................................................
DRCprov <- sf::st_as_sf(readRDS("data/map_bases/gadm/gadm36_COD_1_sp.rds"))
ge <- readRDS("data/derived_data/spacemips_GE.rds")
load("data/map_bases/space_mips_maps_bases.rda")

# read base data
drcsmpls <- readRDS("data/derived_data/sample_metadata.rds") %>% 
  dplyr::select(c("barcode", "hv001"))

```

## Sampling Locations
```{r}

#..............................................................
# wrangle for plot
#..............................................................
drcsmpls.ge <- dplyr::left_join(drcsmpls, ge, by = "hv001")
drcsmpls.ge <- drcsmpls.ge %>% 
  dplyr::group_by(hv001) %>% 
  dplyr::summarise(n = n(),
                   longnum = mean(longnum),
                   latnum = mean(latnum)) # geocoords same w/in clust

drcsmpls.ge$smpls <- "Y"
ge.notsmpl <- ge %>% 
  dplyr::select(c("hv001", "longnum", "latnum")) %>% 
  dplyr::filter(!hv001 %in% drcsmpls.ge$hv001) %>% 
  dplyr::mutate(
    n = 1, 
    smpls = "N"
  )
# bring together for plot
ge.plotpts <- rbind.data.frame(drcsmpls.ge, ge.notsmpl)

#..............................................................
# plot
#..............................................................
smplocations <- ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  prettybasemap_nodrc_nonorth_dark +
  geom_point(data = ge.plotpts, 
             aes(x = longnum, y = latnum, size = n, color = factor(smpls), shape = factor(smpls)), 
             alpha = 0.8) +
  scale_color_manual(name = "DHS Cluster", 
                     labels = c("Not Sampled", "Sampled"),
                     values = c("#a6bddb", "#ef3b2c")) +
  scale_shape_manual(name = "DHS Cluster", 
                     labels = c("Not Sampled", "Sampled"),
                     values = c(4, 20)) +
  scale_size("Cluster Size", range = c(1, 4)) +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold", size = 12, vjust = 0.5, hjust = 0),
    legend.text = element_text(face = "bold", size = 11),
    plot.margin = unit(c(0.05, 0.05, 0, 1.5),"cm"))

jpgsnapshot(outpath = "results/figures/smplocations_clster.jpg",
            plot = smplocations)
```
```{r, results='asis'}
plot(smplocations)
```

## Road Netwok and Airports
```{r}
#.............................. 
# road network
#............................... 
roadnetworkplotObj <- readRDS("data/distance_data/roadnetworkplotObj.RDS") +
  theme(plot.margin = unit(c(0.05, 0.05, 0, 0.05),"cm"))


#.............................. 
# airport locations
#............................... 
airports <- readr::read_csv("data/raw_data/flight_data/hotosm_cd-airports.csv") %>%
  dplyr::filter(type %in% c("large_airport", "medium_airport")) %>%
  dplyr::select(c("name", "longitude_deg", "latitude_deg")) %>% 
  dplyr::mutate(longitude_deg = as.numeric(longitude_deg),
                latitude_deg = as.numeric(latitude_deg))

airportplotObj <- ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  prettybasemap_nodrc_nonorth_dark +
  geom_point(data = airports, aes(x = longitude_deg, y = latitude_deg), 
             color = "#a6bddb") + 
  ggrepel::geom_text_repel(data = airports, 
                           aes(x = longitude_deg, y = latitude_deg, label = name),
                           size = 2.5, color = "#f0f0f0")
plot(airportplotObj)

```


```{r, results='asis'}


# bring together
MainFig <- cowplot::plot_grid(roadnetworkplotObj, 
                              airportplotObj,
                              nrow = 1, ncol = 2, labels = c("(A)", "(B)"))

jpeg("results/figures/drc_dists_combined_collated.jpg", width = 8, height = 11, units = "in", res = 500)
plot(MainFig)
graphics.off()

```
```{r, results='asis'}
plot(MainFig)
```



## Clusters with Single Sample
```{r, results='asis'}

# bring together for plot
ge.plotpts.onesample <- rbind.data.frame(drcsmpls.ge, ge.notsmpl) %>% 
  dplyr::filter(n == 1) %>% 
  dplyr::filter(smpls == "Y")

#..............................................................
# plot
#..............................................................
smplocations.onesample <- ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  prettybasemap_nodrc_nonorth_dark +
  geom_point(data = ge.plotpts.onesample, 
             aes(x = longnum, y = latnum, 
                 alpha = 0.8), color = "#ef3b2c", shape = 8) +
  theme(
    legend.position = "none",
    legend.title = element_text(face = "bold", size = 12, vjust = 0.5, hjust = 0.5),
    legend.text = element_text(face = "bold", size = 11)
  ) 

jpgsnapshot(outpath  = "results/figures/smplocations_one_sample.jpg",
            plot = smplocations.onesample)

```
```{r, results='asis'}
plot(smplocations.onesample)
```
**Overall, there are `r nrow(ge.plotpts.onesample)` clusters that only have a single sample within the cluster.

## Geodistances
```{r}
distancemat <- readRDS("data/distance_data/distancematrix_bycluster.rds")

distmatplot <- distancemat %>% 
  tidyr::pivot_longer(., cols = -c("hv001.x", "hv001.y"),
                      names_to = "distlvl", values_to = "dist") %>% 
  
  dplyr::mutate(dist = dist/1e3,
                hv001.x = forcats::fct_rev(forcats::fct_reorder(.f = hv001.x, 
                                                                .x = hv001.x, 
                                                                .fun = length)),
                hv001.y = forcats::fct_rev(forcats::fct_reorder(.f = hv001.y, 
                                                                .x = hv001.y, 
                                                                .fun = length))) %>%
  dplyr::mutate(distlvl = factor(distlvl, levels = c("gcdistance", "roaddistance", "airportdistance"), labels = c("GC Dist.", "Road Dist.", "Airport Dist."))) %>% 
  ggplot() + 
  geom_tile(aes(x = hv001.x, y = hv001.y, fill = dist)) +
  scale_fill_viridis_c("Distance (km)") + 
  xlab("DHS Cluster") + ylab("DHS Cluster") +
  facet_wrap(. ~ distlvl, shrink=F, scales = "free") +
  theme_minimal() + 
  theme(
    axis.title.y = element_text(vjust = 0.5, hjust = 0.5, size = 11, face = "bold"), 
    axis.title.x = element_text(vjust = 0.5, hjust = 0.5, size = 11, face = "bold"),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "right",
    legend.text = element_text(hjust = 0.5, vjust = 0.5, size = 8),
    plot.background = element_blank(),
    panel.grid = element_blank()
  )

# plot out
jpeg("results/figures/geodistance_matrix_plots.jpg", width = 11, height = 8, units = "in", res = 800)
plot(distmatplot)
graphics.off()
```
```{r, results='asis'}
plot(distmatplot)
```
