---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Sample and Study Site Overview 
Basic descriptive data for samples and study location. 
```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = T, results = 'hide',
                      fig.align = 'center', fig.width = 8, fig.height = 8)
knitr::opts_knit$set(root.dir = here::here())
```
```{r}
library(tidyverse)
library(sf)
source("R/basics.R")
```

## Sample Characteristics
Samples are combination of adults, children with a full DHS line, and children that were "accidentally" sampled (i.e. not a full DHS line -- occurred when child was slightly older than 5 yrs but a DBS was still taken). 
```{r}
# read base data
drcsmpls <- readRDS("data/derived_data/sample_metadata.rds") %>% 
  dplyr::select(c("barcode", "hv001"))
# bring in DHS


```

## Sampling Locations
```{r}
#..............................................................
# import for plotting
#..............................................................
DRCprov <- sf::st_as_sf(readRDS("data/map_bases/gadm/gadm36_COD_1_sp.rds"))
ge <- readRDS("data/derived_data/spacemips_GE.rds")
load("data/map_bases/space_mips_maps_bases.rda")

#..............................................................
# wrangle for plot
#..............................................................
drcsmpls.ge <- dplyr::left_join(drcsmpls, ge, by = "hv001")
drcsmpls.ge <- drcsmpls.ge %>% 
  dplyr::group_by(hv001) %>% 
  dplyr::summarise(n = n(),
                   longnum = mean(longnum),
                   latnum = mean(latnum)) # geocoords same w/in clust

drcsmpls.ge$smpls <- "Y"
ge.notsmpl <- ge %>% 
  dplyr::select(c("hv001", "longnum", "latnum")) %>% 
  dplyr::filter(!hv001 %in% drcsmpls.ge$hv001) %>% 
  dplyr::mutate(
    n = 1, 
    smpls = "N"
  )
# bring together for plot
ge.plotpts <- rbind.data.frame(drcsmpls.ge, ge.notsmpl)

#..............................................................
# plot
#..............................................................
smplocations <- ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  prettybasemap_nodrc_dark +
  geom_point(data = ge.plotpts, 
             aes(x = longnum, y = latnum, size = n, color = factor(smpls), shape = factor(smpls)), 
             alpha = 0.8) +
  scale_color_manual(name = "DHS Cluster", 
                     labels = c("Not Sampled", "Sampled"),
                     values = c("#a6bddb", "#ef3b2c")) +
  scale_shape_manual(name = "DHS Cluster", 
                     labels = c("Not Sampled", "Sampled"),
                     values = c(4, 20)) +
  scale_size("Cluster Size", range = c(1, 4)) +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold", size = 12, vjust = 0.5, hjust = 0),
    legend.text = element_text(face = "bold", size = 11)
  ) 
jpgsnapshot(outpath = "results/figures/smplocations_clster.jpg",
            plot = smplocations)
```
```{r, results='asis'}
plot(smplocations)
```
```{r, results='asis'}
roadnetworkplotObj <- readRDS("data/distance_data/roadnetworkplotObj.RDS")
rivernetworkplotObj <- readRDS("data/distance_data/rivernetworkplotObj.RDS")


cowplot::plot_grid(smplocations, roadnetworkplotObj, rivernetworkplotObj,
                   nrow = 1, align = "h")
```

```{r}

smplocations.nolegend <- ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  prettybasemap_nodrc_dark +
  geom_point(data = ge.plotpts, 
             aes(x = longnum, y = latnum, size = n, color = factor(smpls), shape = factor(smpls)), 
             alpha = 0.8) +
  scale_color_manual("DHS \n Cluster", values = c("#a6bddb", "#ef3b2c")) +
  scale_size("Cluster \n Size", range = c(1, 4)) +
  theme(
    legend.position = "none",
    legend.title = element_text(face = "bold", size = 12, vjust = 0.5, hjust = 0.5),
    legend.text = element_text(face = "bold", size = 11)
  ) 

distnetplot <- cowplot::plot_grid(smplocations.nolegend, roadnetworkplotObj, rivernetworkplotObj,
                                  nrow = 1, align = "h", labels = c("(A)", "(B)", "(C)"))

jpgsnapshot(outpath  = "results/figures/smplocationsroad_river_clster.jpg",
            plot = distnetplot)

```

## Clusters with Single Sample
```{r}

# bring together for plot
ge.plotpts.onesample <- rbind.data.frame(drcsmpls.ge, ge.notsmpl) %>% 
  dplyr::filter(n == 1) %>% 
  dplyr::filter(smpls == "Y")

#..............................................................
# plot
#..............................................................
smplocations.onesample <- ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  prettybasemap_nodrc_dark +
  geom_point(data = ge.plotpts.onesample, 
             aes(x = longnum, y = latnum, 
                 alpha = 0.8), color = "#ef3b2c", shape = 8) +
  theme(
    legend.position = "none",
    legend.title = element_text(face = "bold", size = 12, vjust = 0.5, hjust = 0.5),
    legend.text = element_text(face = "bold", size = 11)
  ) 

jpgsnapshot(outpath  = "results/figures/smplocations_one_sample.jpg",
            plot = smplocations.onesample)

```
```{r, results='asis'}
plot(smplocations.onesample)
```
**Overall, there are `r nrow(ge.plotpts.onesample)` clusters that only have a single sample within the cluster.


