---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Genetic Loci Descriptive Statistics 

```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = T, results = 'hide',
                      fig.align = 'center', fig.width = 8, fig.height = 8)
knitr::opts_knit$set(root.dir = here::here())
```


```{r}
source("R/basics.R")
source("R/pairwise_helpers.R")
library(tidyverse)
library(cowplot)
#...................... 
# metadata
#......................
mtdt <- readRDS("data/derived_data/sample_metadata.rds") %>% 
  magrittr::set_colnames(tolower(colnames(.))) %>%
  dplyr::select(c("name", "country", "hv001", "adm1name", "longnum", "latnum"))

#......................
# Import Genetic Data
#......................
ibD <- readRDS("data/derived_data/bigbarcode_genetic_data/mipanalyzer.DRCibD.long.mtdt.rds")
ibS <- readRDS("data/derived_data/bigbarcode_genetic_data/mipanalyzer.DRCibS.long.mtdt.rds")
vcf.DRC <- vcfR::read.vcfR("data/derived_data/bigbarcode_genetic_data/mipbivcfR.DRC.vcf")
CHROMPOS <- tibble::tibble(CHROM = vcfR::getCHROM(vcf.DRC),
                           POS = vcfR::getPOS(vcf.DRC))
```

## Overview of Loci Genetic Information
### Number of sites 

```{r, results='asis'}
CHROMPOS.num <- CHROMPOS %>% 
  dplyr::group_by(CHROM) %>% 
  dplyr::summarise(n = n())

DT::datatable(CHROMPOS.num, extensions='Buttons',
              options = list(
                searching = T,
                pageLength = 14,
                dom = 'Bfrtip', 
                buttons = c('csv')))

```


## Pairwise Genetics Comparions
Let's look at pairwise comparison of the identity by state and identity by descent of all of our samples. 
### Pairwise IBS
```{r}
# Pairwise Comparisons -- want full matrix
ibS.long.smpl <- ibS %>% 
  dplyr::select(c("smpl1", "smpl2", "hammings")) %>% 
  expand_distance_matrix(.)

ibS.long.clst <- ibS %>% 
  dplyr::select(c("hv001.x", "hv001.y", "hammings")) %>% 
  expand_distance_matrix(.)

ibS.long.prov <- ibS %>% 
  dplyr::select(c("adm1name.x", "adm1name.y", "hammings")) %>% 
  expand_distance_matrix(.)
```

###### Sample
```{r, results='asis'}
ibS.long.smpl %>% # selfs already excluded
  summ_var(df = ., x = "hammings", groupingvar = "smpl1") %>% 
  pretty_DT_tab(.)

```

##### Cluster
```{r, results='asis'}
ibS.long.clst %>% # selfs already excluded
  dplyr::group_by(hv001.x) %>% 
  summ_var(df = ., x = "hammings", groupingvar = "hv001.x") %>% 
  pretty_DT_tab(.)
```

##### Prov
```{r, results='asis'}
ibS.long.prov %>% # selfs already excluded
  summ_var(df = ., x = "hammings", groupingvar = "adm1name.x") %>% 
  pretty_DT_tab(.)

```

```{r, results='asis', fig.width = 11, fig.height = 8, fig.align = 'center'}
#...................... 
# plot this
#......................
# smpl
smpl_plot <- ibS.long.smpl %>%
  dplyr::mutate(smpl1 = factor(smpl1),
                smpl2 = factor(smpl2)) %>%
  ggplot() +
  geom_tile(aes(x=smpl1, y=smpl2, fill = hammings)) +
  xlab("Samples") + ylab("Samples") + 
  scale_fill_viridis_c("Mean IBS") +
  theme_minimal() +
  theme(
    axis.title = element_text(vjust = 0.5, hjust = 0.5, size = 11, face = "bold"),
    axis.text = element_blank(),
    legend.position = "none",
    plot.background = element_blank(),
    panel.grid = element_blank()
  )
# cluster -- white spaces are those 93 clusters with only one sample, no w/in comparison
clst_plot <- ibS.long.clst %>%
  dplyr::group_by_at(c("hv001.x", "hv001.y"), .drop = F) %>% 
  dplyr::summarise(
    meanIBS = mean(hammings)
  ) %>% 
  dplyr::mutate(hv001.x = factor(hv001.x),
                hv001.y = factor(hv001.y)) %>%
  ggplot() +
  geom_tile(aes(x=hv001.x, y=hv001.y, fill = meanIBS)) +
  xlab("Clusters") + ylab("Clusters") + 
  scale_fill_viridis_c("IBS") +
  theme_minimal() +
  theme(
    axis.title = element_text(vjust = 0.5, hjust = 0.5, size = 11, face = "bold"),
    axis.text = element_blank(),
    legend.position = "none",
    plot.background = element_blank(),
    panel.grid = element_blank()
  )

# prov
prov_plot <- ibS.long.prov %>%
  dplyr::group_by_at(c("adm1name.x", "adm1name.y")) %>% 
  dplyr::summarise(
    meanIBS = mean(hammings)
  ) %>% 
  dplyr::mutate(adm1name.x = factor(adm1name.x),
                adm1name.y = factor(adm1name.y)) %>%
  ggplot() +
  geom_tile(aes(x=adm1name.x, y=adm1name.y, fill = meanIBS)) +
  xlab("Provinces") + ylab("Provinces") + 
  scale_fill_viridis_c("IBS") +
  theme_minimal() +
  theme(
    axis.title = element_text(vjust = 0.5, hjust = 0.5, size = 11, face = "bold"),
    axis.text = element_blank(),
    legend.position = "none",
    plot.background = element_blank(),
    panel.grid = element_blank()
  )


legend <- get_legend(
  clst_plot +  
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom",
          legend.title = element_text(vjust = 0.5, hjust = 0.5, size = 14, face = "bold"),
          legend.text = element_text(vjust = 0.5, hjust = 0.5, size = 12, face = "bold")
    )
)
firstrow <- cowplot::plot_grid(smpl_plot, clst_plot, prov_plot, 
                               labels = c("(A)", "(B)", "(C)"),
                               align = "h")

cowplot::plot_grid(firstrow, legend, ncol = 1, rel_heights = c(1, 0.1))



```



### Pairwise IBD
```{r}
# Pairwise Comparisons -- want full matrix
ibD.long.smpl <- ibD %>% 
  dplyr::select(c("smpl1", "smpl2", "malecotf")) %>% 
  expand_distance_matrix(.)

ibD.long.clst <- ibD %>% 
  dplyr::select(c("hv001.x", "hv001.y", "malecotf")) %>% 
  expand_distance_matrix(.)

ibD.long.prov <- ibD %>% 
  dplyr::select(c("adm1name.x", "adm1name.y", "malecotf")) %>% 
  expand_distance_matrix(.)
```


###### Sample
```{r, results='asis'}
ibD.long.smpl %>% # selfs already excluded
  summ_var(df = ., x = "malecotf", groupingvar = "smpl1") %>% 
  pretty_DT_tab(.)

```


##### Cluster
```{r, results='asis'}
ibD.long.clst %>% # selfs already excluded
  summ_var(df = ., x = "malecotf", groupingvar = "hv001.x") %>% 
  pretty_DT_tab(.)

```

##### Prov
```{r, results='asis'}
ibD.long.prov %>% # selfs already excluded
  summ_var(df = ., x = "malecotf", groupingvar = "adm1name.x") %>% 
  pretty_DT_tab(.)
```

```{r, results='asis', fig.width = 11, fig.height = 8, fig.align = 'center'}
#...................... 
# plot this
#......................
# smpl
smpl_plot <- ibD.long.smpl %>%
  dplyr::mutate(smpl1 = factor(smpl1),
                smpl2 = factor(smpl2)) %>%
  ggplot() +
  geom_tile(aes(x=smpl1, y=smpl2, fill = malecotf)) +
  xlab("Samples") + ylab("Samples") + 
  scale_fill_viridis_c("Mean IBD") +
  theme_minimal() +
  theme(
    axis.title = element_text(vjust = 0.5, hjust = 0.5, size = 11, face = "bold"),
    axis.text = element_blank(),
    legend.position = "none",
    plot.background = element_blank(),
    panel.grid = element_blank()
  )
# cluster -- white spaces are those 93 clusters with only one sample, no w/in comparison
clst_plot <- ibD.long.clst %>%
  dplyr::group_by_at(c("hv001.x", "hv001.y"), .drop = F) %>% 
  dplyr::summarise(
    meanIBD = mean(malecotf)
  ) %>% 
  dplyr::mutate(hv001.x = factor(hv001.x),
                hv001.y = factor(hv001.y)) %>%
  ggplot() +
  geom_tile(aes(x=hv001.x, y=hv001.y, fill = meanIBD)) +
  xlab("Clusters") + ylab("Clusters") + 
  scale_fill_viridis_c("Mean IBD") +
  theme_minimal() +
  theme(
    axis.title = element_text(vjust = 0.5, hjust = 0.5, size = 11, face = "bold"),
    axis.text = element_blank(),
    legend.position = "none",
    plot.background = element_blank(),
    panel.grid = element_blank()
  )

# prov
prov_plot <- ibD.long.prov %>%
  dplyr::group_by_at(c("adm1name.x", "adm1name.y")) %>% 
  dplyr::summarise(
    meanIBD = mean(malecotf)
  ) %>% 
  dplyr::mutate(adm1name.x = factor(adm1name.x),
                adm1name.y = factor(adm1name.y)) %>%
  ggplot() +
  geom_tile(aes(x=adm1name.x, y=adm1name.y, fill = meanIBD)) +
  xlab("Provinces") + ylab("Provinces") + 
  scale_fill_viridis_c("Mean IBD") +
  theme_minimal() +
  theme(
    axis.title = element_text(vjust = 0.5, hjust = 0.5, size = 11, face = "bold"),
    axis.text = element_blank(),
    legend.position = "none",
    plot.background = element_blank(),
    panel.grid = element_blank()
  )


legend <- get_legend(
  clst_plot +  
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom",
          legend.title = element_text(vjust = 0.5, hjust = 0.5, size = 14, face = "bold"),
          legend.text = element_text(vjust = 0.5, hjust = 0.5, size = 12, face = "bold")
    )
)
firstrow <- cowplot::plot_grid(smpl_plot, clst_plot, prov_plot, 
                               labels = c("(A)", "(B)", "(C)"),
                               align = "h")

cowplot::plot_grid(firstrow, legend, ncol = 1, rel_heights = c(1, 0.1))



```



### Distribution of Relatedness
When considering pairwise IBD as the outcome, we can use a log-base-2 tranformation to normalize the IBD measure. Log-base-2 represents the number of generations ago to samples were IBD (or coalesced) under an assumption of outcrossing. _N.B. larger numbers mean more generations (less related)._

#### All Samples
```{r, results='asis'}
ibD.exp.hv001 %>% 
  dplyr::mutate(malecotf_gens = -log2(malecotf),
                malecotf_gens_inv = 1/malecotf_gens) %>% 
  dplyr::filter(malecotf_gens != Inf) %>% 
  ggplot() + 
  geom_histogram(aes(x = malecotf_gens, y = ..count.. )) + 
  plot_theme +
  xlab("Generations") + ylab("")

```

#### Prov
```{r, results='asis'}

p1 <- mtdt %>% 
  dplyr::select(c("name", "adm1name")) %>% 
  dplyr::rename(smpl1 = name)

p2 <- mtdt %>% 
  dplyr::select(c("name", "adm1name")) %>% 
  dplyr::rename(smpl2 = name)

ibD.exp.hv001 %>% 
  dplyr::left_join(., y = p1, by = "smpl1") %>% 
  dplyr::left_join(., y = p2, by = "smpl2") %>% 
  expand_distance_matrix() %>% 
  dplyr::filter(malecotf_gens != Inf) %>% 
  ggplot() + 
  ggridges::stat_density_ridges(
    geom = "density_ridges_gradient", calc_ecdf = T,
    quantiles = 4, quantile_lines = T, 
    aes(x = malecotf_gens, y = adm1name.x, fill = factor(..quantile..))
  ) +
  scale_fill_viridis_d("Quantiles") +
  ylab("Province") + xlab("Distribution of Generations") +
  plot_theme 

```
