---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = T, results = 'hide',
                      fig.align = 'center', fig.width = 8, fig.height = 8)
knitr::opts_knit$set(root.dir = here::here())
```

```{r}
#---------------------------------------------------------------------------
# Purpose of this script is to look
# at decay of IBD with distance
#----------------------------------------------------------------------------
source("R/pairwise_helpers.R")
source("R/themes.R")
library(tidyverse)

#....................................................................................
# Import Genetic Data
#....................................................................................
ibD <- readRDS("data/derived_data/bigbarcode_genetic_data/mipanalyzer.DRCibD.long.mtdt.rds") %>% 
  dplyr::mutate(hv001.x = as.character(hv001.x),
                hv001.y = as.character(hv001.y))
ibS <- readRDS("data/derived_data/bigbarcode_genetic_data/mipanalyzer.DRCibS.long.mtdt.rds") %>% 
  dplyr::mutate(hv001.x = as.character(hv001.x),
                hv001.y = as.character(hv001.y))

```

```{r}
#.......................................................................
# Import Distance matrix
#....................................................................... 
distancematrix.cluster <- readRDS("data/distance_data/distancematrix_bycluster.rds")


#.......................................................................
# Merge Geo and Genetic
#....................................................................... 

# need to make the distance matrices "wide"
ibSdist <- long_distance_matrix_join(x=ibS, y=distancematrix.cluster,
                                     by = c("hv001.x", "hv001.y")) %>%
  dplyr::mutate(gcdistance = gcdistance/1e3,
                roaddistance = roaddistance/1e3,
                airportdistance = airportdistance/1e3,
                airportdistance = as.numeric(airportdistance)) # remove units
# take care of cluster diagnonals
ibSdist$gcdistance[ibSdist$hv001.x == ibSdist$hv001.y] <- 0
ibSdist$roaddistance[ibSdist$hv001.x == ibSdist$hv001.y] <- 0
ibSdist$airportdistance[ibSdist$hv001.x == ibSdist$hv001.y] <- 0


ibDdist <- long_distance_matrix_join(x=ibD, y=distancematrix.cluster,
                                     by = c("hv001.x", "hv001.y")) %>%
  dplyr::mutate(gcdistance = gcdistance/1e3,
                roaddistance = roaddistance/1e3,
                airportdistance = airportdistance/1e3,
                airportdistance = as.numeric(airportdistance)) # remove units


# take care of cluster diagnonals
ibDdist$gcdistance[ibDdist$hv001.x == ibDdist$hv001.y] <- 0
ibDdist$roaddistance[ibDdist$hv001.x == ibDdist$hv001.y] <- 0
ibDdist$airportdistance[ibDdist$hv001.x == ibDdist$hv001.y] <- 0



```



# Replicating Figure 4 from [Verity et al. 2019](https://www.biorxiv.org/content/10.1101/656561v1)
```{r}

mainplot <- ibDdist %>% 
ggplot() + 
  geom_histogram(aes(x=malecotf, y = (..count../sum(..count..))*100), 
                 color = "#000000", fill = "#d9d9d9") + 
  xlab("IBD") + ylab("frequency (%)") + 
  theme_classic()

insetplot <- ibDdist %>% 
ggplot() + 
  geom_histogram(aes(x=malecotf, y = (..count../sum(..count..))*100), 
                 color = "#000000", fill = "#d9d9d9") + 
  xlab("IBD") + ylab("frequency (%)") + 
  theme_classic() + 
  coord_cartesian(xlim = c(0.5,1), ylim = c(0,0.002)) +
  theme_bw() + 
  theme(panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent"))

```

```{r, results='asis'}

cowplot::ggdraw() +
  cowplot::draw_plot(mainplot, x = 0, y = 0, width = 1, height = 1, scale = 1) +
  cowplot::draw_plot(insetplot, x = 0.5, y= 0.3, width = 0.4, height = 0.4)


```



## IBS vs. Distances

### IBS - Greater Circle
```{r}
ibSdist$gcdistance_cat <- cut(x = c(ibSdist$gcdistance),
                                breaks = c(0, 1e-26,
                                           seq(100, 2000, by = 100),
                                           Inf),
                                right = F, #
                                labels = c("Within", "100km",
                                           "200km", "300km", "400km", "500km", "600km",
                                           "700km", "800km", "900km", "1000km", "1100km",
                                           "1200km", "1300km", "1400km", "1500km", "1600km",
                                           "1700km", "1800km", "1900km", "2000km", ">2000km"
                                           )
                                )



```

```{r, results='asis'}
ibSdist.gc.summary <- ibSdist %>%
  dplyr::group_by(gcdistance_cat) %>%
  dplyr::summarise(
    n = n(),
    IBSmin = min(hammings),
    IBS25 = quantile(hammings, c(0.25)),
    IBSmean = mean(hammings),
    IBSsd = sd(hammings),
    IBSmedian = median(hammings),
    IBS75 = quantile(hammings, c(0.75)),
    IBSmax = max(hammings)
  )


ibSdist.gc.summary %>% 
  dplyr::mutate_if(is.numeric, round, 2) %>% 
  DT::datatable(., extensions='Buttons',
               options = list(
              searching = T,
              pageLength = 20,
              dom = 'Bfrtip', 
              buttons = c('csv')))

```



### IBS - Road Distance
```{r}
ibSdist$roaddistance_cat <- cut(x = c(ibSdist$roaddistance),
                                breaks = c(0, 1e-26, 
                                           seq(100, 2000, by = 100),
                                           Inf),
                                right = F, #
                                labels = c("Within", "100km",
                                           "200km", "300km", "400km", "500km", "600km",
                                           "700km", "800km", "900km", "1000km", "1100km",
                                           "1200km", "1300km", "1400km", "1500km", "1600km",
                                           "1700km", "1800km", "1900km", "2000km", ">2000km"
                                           )
                                )

```

```{r, results='asis'}
ibSdist.road.summary <- ibSdist %>%
  dplyr::group_by(roaddistance_cat) %>%
  dplyr::summarise(
    n = n(),
    IBSmin = min(hammings),
    IBS25 = quantile(hammings, c(0.25)),
    IBSmean = mean(hammings),
    IBSsd = sd(hammings),
    IBSmedian = median(hammings),
    IBS75 = quantile(hammings, c(0.75)),
    IBSmax = max(hammings)
  )


ibSdist.road.summary %>% 
  dplyr::mutate_if(is.numeric, round, 2) %>% 
  DT::datatable(., extensions='Buttons',
               options = list(
              searching = T,
              pageLength = 20,
              dom = 'Bfrtip', 
              buttons = c('csv')))

```





### IBS - Airport Distance
```{r}
ibSdist$airportdistance_cat <- cut(x = c(ibSdist$airportdistance),
                                breaks = c(0, 1e-26,
                                           seq(100, 2000, by = 100),
                                           Inf),
                                right = F, #
                                labels = c("Within", "100km",
                                           "200km", "300km", "400km", "500km", "600km",
                                           "700km", "800km", "900km", "1000km", "1100km",
                                           "1200km", "1300km", "1400km", "1500km", "1600km",
                                           "1700km", "1800km", "1900km", "2000km", ">2000km"
                                           )
                                )

```

```{r, results='asis'}

ibSdist.airport.summary <- ibSdist %>%
  dplyr::group_by(airportdistance_cat) %>%
  dplyr::summarise(
    n = n(),
    IBSmin = min(hammings),
    IBS25 = quantile(hammings, c(0.25)),
    IBSmean = mean(hammings),
    IBSsd = sd(hammings),
    IBSmedian = median(hammings),
    IBS75 = quantile(hammings, c(0.75)),
    IBSmax = max(hammings)
  )


ibSdist.airport.summary %>% 
  dplyr::mutate_if(is.numeric, round, 2) %>% 
  DT::datatable(., extensions='Buttons',
               options = list(
              searching = T,
              pageLength = 20,
              dom = 'Bfrtip', 
              buttons = c('csv')))

```

## IBS Decay Plots - Continuous

```{r}
IBSdecay.df <- ibSdist %>%
  dplyr::select(c("smpl1", "smpl2", "hammings", dplyr::ends_with("_cat"))) %>%
  tidyr::gather(., key = "distcat", value = "distmetric", 4:6) %>%
  dplyr::mutate(distcat = factor(distcat, levels = c("gcdistance_cat", "roaddistance_cat",
                                                     "airportdistance_cat"), labels = c("GC Dist.", "Road Dist.", "Airport Dist.")),
                distmetric = factor(distmetric, levels = c("Within", "100km",
                                           "200km", "300km", "400km", "500km", "600km",
                                           "700km", "800km", "900km", "1000km", "1100km",
                                           "1200km", "1300km", "1400km", "1500km", "1600km",
                                           "1700km", "1800km", "1900km", "2000km", ">2000km"
                                           )))

IBSdecay.plot <- IBSdecay.df %>%
  ggplot() +
  geom_boxplot(aes(x=distmetric, y = hammings)) +
  facet_wrap(distcat~., scales = "free") +
  plot_theme +
  theme(axis.text.x = element_text(angle = 90))


jpeg("results/figures/IBS_decay_plot_continuous.jpg",
     unit = "in", height = 8, width = 11, res = 500)
plot(IBSdecay.plot)
graphics.off()

```

```{r, results='asis'}
plot(IBSdecay.plot)

```

## IBS Decay Plots - Means

```{r}

IBSdecay.df <- ibSdist %>%
  dplyr::select(c("smpl1", "smpl2", "hammings", dplyr::ends_with("distance"), "airportdistance")) %>%
  tidyr::gather(., key = "distcat", value = "distmetric", 4:6) %>%
  dplyr::mutate(distmetcat = cut(x = distmetric, breaks = c(0, 1e-26,  
                                                            seq(100, 2000, by = 100),
                                                            Inf),
                                 right = F, #
                                 labels = c("Within", "100km",
                                            "200km", "300km", "400km", "500km", "600km",
                                            "700km", "800km", "900km", "1000km", "1100km",
                                            "1200km", "1300km", "1400km", "1500km", "1600km",
                                            "1700km", "1800km", "1900km", "2000km", ">2000km")
                                 ),
                distcat = factor(distcat, levels = c("gcdistance", "roaddistance",
                                                    "airportdistance"), 
                                 labels = c("GC Dist.", "Road Dist.", "Airport Dist."))
                ) %>% 
  dplyr::group_by(distcat, distmetcat) %>% 
  dplyr::summarise(
    n = n(),
    meanIBS = mean(hammings),
    sdIBS = sd(hammings),
    seIBS = sdIBS/sqrt(n),
    U95CI = meanIBS + 1.96 * seIBS,
    L95CI = meanIBS - 1.96 * seIBS
  )
  
  
  
IBSdecay.plot <- IBSdecay.df %>%
  ggplot() +
  geom_pointrange(aes(x = distmetcat, y = meanIBS, ymin = L95CI, ymax = U95CI)) + 
  scale_size(range = c(0, 1)) +
  scale_y_continuous("Mean IBS") +
  facet_wrap(distcat~.) +
  plot_theme +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        axis.text.x = element_text(angle = 90, size = 8, face = "bold"),
        axis.line.x = element_line(size = 1.1, color = "#000000"),
        panel.grid.minor.x = element_line(color = "#bdbdbd"),
        panel.grid.minor.y = element_line(color = "#bdbdbd")
        )


jpeg("results/figures/IBS_decay_plot_meanIBS.jpg",
     unit = "in", height = 8, width = 11, res = 500)
plot(IBSdecay.plot)
graphics.off()

```

```{r, results='asis'}
plot(IBSdecay.plot)

```



## IBS Decay by Province
Note, here we are including all pairwise comparisons for a given sample in a province. 
### IBS Decay Plots - Continuous, Facet Province

```{r}

ibSdist.long <- expand_distance_matrix(ibSdist) %>% 
  dplyr::rename(adm1name = adm1name.x)


IBSdecay.df <- ibSdist.long %>%
  dplyr::select(c("smpl1", "smpl2", "adm1name", "hammings", dplyr::ends_with("_cat"))) %>%
  tidyr::gather(., key = "distcat", value = "distmetric", 5:7) %>%
  dplyr::mutate(distcat = factor(distcat, levels = c("gcdistance_cat", "roaddistance_cat",
                                                     "airportdistance_cat"), labels = c("GC Dist.", "Road Dist.", "Airport Dist.")),
                distmetric = factor(distmetric, levels = c("Within", "100km",
                                           "200km", "300km", "400km", "500km", "600km",
                                           "700km", "800km", "900km", "1000km", "1100km",
                                           "1200km", "1300km", "1400km", "1500km", "1600km",
                                           "1700km", "1800km", "1900km", "2000km", ">2000km"
                                           )))

IBSdecay.plot <- IBSdecay.df %>%
  ggplot() +
  geom_boxplot(aes(x=distmetric, y = hammings)) +
  facet_grid(adm1name~distcat, scales = "free") +
  plot_theme +
  theme(axis.text.x = element_text(angle = 90))


jpeg("results/figures/IBS_decay_plot_continuous_adm1facet.jpg",
     unit = "in", height = 25, width = 11, res = 500)
plot(IBSdecay.plot)
graphics.off()

```

```{r, results='asis', fig.height=25, fig.width=11}
plot(IBSdecay.plot)

```


### IBS Decay Plots - Means, Facet Province

```{r}


IBSdecay.df <- ibSdist.long %>%
    dplyr::select(c("smpl1", "smpl2", "adm1name", "hammings", dplyr::ends_with("distance"), "airportdistance")) %>%
  tidyr::gather(., key = "distcat", value = "distmetric", 5:7) %>%
  dplyr::mutate(distmetcat = cut(x = distmetric, breaks = c(0, 1e-26,  
                                                            seq(100, 2000, by = 100),
                                                            Inf),
                                 right = F, #
                                 labels = c("Within", "100km",
                                            "200km", "300km", "400km", "500km", "600km",
                                            "700km", "800km", "900km", "1000km", "1100km",
                                            "1200km", "1300km", "1400km", "1500km", "1600km",
                                            "1700km", "1800km", "1900km", "2000km", ">2000km")
                                 ),
                distcat = factor(distcat, levels = c("gcdistance", "roaddistance",
                                                    "airportdistance"), 
                                 labels = c("GC Dist.", "Road Dist.", "Airport Dist."))
                ) %>% 
  dplyr::group_by(adm1name, distcat, distmetcat) %>% 
  dplyr::summarise(
    n = n(),
    meanIBS = mean(hammings),
    sdIBS = sd(hammings),
    seIBS = sdIBS/sqrt(n),
    U95CI = meanIBS + 1.96 * seIBS,
    L95CI = meanIBS - 1.96 * seIBS
  )
  

  

IBSdecay.plot <- IBSdecay.df %>%
  ggplot() +
  geom_pointrange(aes(x = distmetcat, y = meanIBS, ymin = L95CI, ymax = U95CI)) + 
  scale_size(range = c(0, 1)) +
  # geom_smooth(aes(x = distmetcat, y = meanIBS), 
  #             method="glm",
  #             formula = y+.Machine$double.xmin ~ x,
  #             method.args=list(family=gaussian(link="log")), color = "#9ecae1") +
  scale_y_continuous("Mean IBS") +
  facet_wrap(adm1name ~ distcat) +
  plot_theme +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        axis.text.x = element_text(angle = 90, size = 8, face = "bold"),
        axis.line.x = element_line(size = 1.1, color = "#000000"),
        panel.grid.minor.x = element_line(color = "#bdbdbd"),
        panel.grid.minor.y = element_line(color = "#bdbdbd")
        )


jpeg("results/figures/IBS_decay_plot_meanIBS_adm1.jpg",
     unit = "in", height = 25, width = 11, res = 500)
plot(IBSdecay.plot)
graphics.off()

```

```{r, results='asis', fig.height=25, fig.width=11}
plot(IBSdecay.plot)

```




## IBD vs. Distances

### IBD - Greater Circle
```{r}
ibDdist$gcdistance_cat <- cut(x = c(ibDdist$gcdistance),
                              breaks = c(0, 1e-26,
                                           seq(100, 2000, by = 100),
                                           Inf),
                                right = F, #
                                labels = c("Within", "100km",
                                           "200km", "300km", "400km", "500km", "600km",
                                           "700km", "800km", "900km", "1000km", "1100km",
                                           "1200km", "1300km", "1400km", "1500km", "1600km",
                                           "1700km", "1800km", "1900km", "2000km", ">2000km"
                                           )
                                )


```

```{r, results='asis'}
ibDdist.gc.summary <- ibDdist %>%
  dplyr::group_by(gcdistance_cat) %>%
  dplyr::summarise(
    n = n(),
    IBDmin = min(malecotf),
    IBD25 = quantile(malecotf, c(0.25)),
    IBDmean = mean(malecotf),
    IBDsd = sd(malecotf),
    IBDmedian = median(malecotf),
    IBD75 = quantile(malecotf, c(0.75)),
    IBDmax = max(malecotf)
  )


ibDdist.gc.summary %>% 
  dplyr::mutate_if(is.numeric, round, 2) %>% 
  DT::datatable(., extensions='Buttons',
                options = list(
                  searching = T,
                  pageLength = 20,
                  dom = 'Bfrtip', 
                  buttons = c('csv')))

```



### IBD - Road Distance
```{r}
ibDdist$roaddistance_cat <- cut(x = c(ibDdist$roaddistance),
                                breaks = c(0, 1e-26,
                                           seq(100, 2000, by = 100),
                                           Inf),
                                right = F, #
                                labels = c("Within", "100km",
                                           "200km", "300km", "400km", "500km", "600km",
                                           "700km", "800km", "900km", "1000km", "1100km",
                                           "1200km", "1300km", "1400km", "1500km", "1600km",
                                           "1700km", "1800km", "1900km", "2000km", ">2000km"
                                           )
                                )


```

```{r, results='asis'}
ibDdist.road.summary <- ibDdist %>%
  dplyr::group_by(roaddistance_cat) %>%
  dplyr::summarise(
    n = n(),
    IBDmin = min(malecotf),
    IBD25 = quantile(malecotf, c(0.25)),
    IBDmean = mean(malecotf),
    IBDsd = sd(malecotf),
    IBDmedian = median(malecotf),
    IBD75 = quantile(malecotf, c(0.75)),
    IBDmax = max(malecotf)
  )


ibDdist.road.summary %>% 
  dplyr::mutate_if(is.numeric, round, 2) %>% 
  DT::datatable(., extensions='Buttons',
                options = list(
                  searching = T,
                  pageLength = 20,
                  dom = 'Bfrtip', 
                  buttons = c('csv')))

```


### IBD - Airport Distance
```{r}
ibDdist$airportdistance_cat <- cut(x = c(ibDdist$airportdistance),
                             breaks = c(0, 1e-26,
                                           seq(100, 2000, by = 100),
                                           Inf),
                                right = F, #
                                labels = c("Within", "100km",
                                           "200km", "300km", "400km", "500km", "600km",
                                           "700km", "800km", "900km", "1000km", "1100km",
                                           "1200km", "1300km", "1400km", "1500km", "1600km",
                                           "1700km", "1800km", "1900km", "2000km", ">2000km"
                                           )
                                )


```

```{r, results='asis'}

ibDdist.airport.summary <- ibDdist %>%
  dplyr::group_by(airportdistance_cat) %>%
  dplyr::summarise(
    n = n(),
    IBDmin = min(malecotf),
    IBD25 = quantile(malecotf, c(0.25)),
    IBDmean = mean(malecotf),
    IBDsd = sd(malecotf),
    IBDmedian = median(malecotf),
    IBD75 = quantile(malecotf, c(0.75)),
    IBDmax = max(malecotf)
  )


ibDdist.airport.summary %>% 
  dplyr::mutate_if(is.numeric, round, 2) %>% 
  DT::datatable(., extensions='Buttons',
                options = list(
                  searching = T,
                  pageLength = 20,
                  dom = 'Bfrtip', 
                  buttons = c('csv')))

```

## IBD Decay Plots - Continuous

```{r}

IBDdecay.df <- ibDdist %>%
  dplyr::select(c("smpl1", "smpl2", "malecotf", dplyr::ends_with("_cat"))) %>%
  tidyr::gather(., key = "distcat", value = "distmetric", 4:6) %>%
  dplyr::mutate(distcat = factor(distcat, levels = c("gcdistance_cat", "roaddistance_cat",
                                                     "airportdistance_cat"), labels = c("GC Dist.", "Road Dist.", "Airport Dist.")),
                distmetric = factor(distmetric, levels = c("Within", "100km",
                                           "200km", "300km", "400km", "500km", "600km",
                                           "700km", "800km", "900km", "1000km", "1100km",
                                           "1200km", "1300km", "1400km", "1500km", "1600km",
                                           "1700km", "1800km", "1900km", "2000km", ">2000km"
                                           )))

IBDdecay.plot <- IBDdecay.df %>%
  ggplot() +
  geom_boxplot(aes(x=distmetric, y = malecotf)) +
  facet_wrap(distcat~., scales = "free") +
  plot_theme +
  theme(axis.text.x = element_text(angle = 90))


jpeg("results/figures/IBD_decay_plot.jpg",
     unit = "in", height = 8, width = 11, res = 500)
plot(IBDdecay.plot)
graphics.off()

```

```{r, results='asis'}
plot(IBDdecay.plot)

```


## IBD Decay Plots - Means

```{r}

IBDdecay.df <- ibDdist %>%
  dplyr::select(c("smpl1", "smpl2", "malecotf", dplyr::ends_with("distance"), "airportdistance")) %>%
  tidyr::gather(., key = "distcat", value = "distmetric", 4:6) %>%
  dplyr::mutate(distmetcat = cut(x = distmetric, breaks = c(0, 1e-26, 
                                                            seq(100, 2000, by = 100),
                                                            Inf),
                                 right = F, #
                                 labels = c("Within", "100km",
                                            "200km", "300km", "400km", "500km", "600km",
                                            "700km", "800km", "900km", "1000km", "1100km",
                                            "1200km", "1300km", "1400km", "1500km", "1600km",
                                            "1700km", "1800km", "1900km", "2000km", ">2000km")
                                 ),
                distcat = factor(distcat, levels = c("gcdistance", "roaddistance",
                                                    "airportdistance"), 
                                 labels = c("GC Dist.", "Road Dist.", "Airport Dist."))
                ) %>% 
  dplyr::group_by(distcat, distmetcat) %>% 
  dplyr::summarise(
    n = n(),
    meanIBD = mean(malecotf),
    sdIBD = sd(malecotf),
    seIBD = sdIBD/sqrt(n),
    U95CI = meanIBD + 1.96 * seIBD,
    L95CI = meanIBD - 1.96 * seIBD
  )
  
  
  

IBDdecay.plot <- IBDdecay.df %>%
  ggplot() +
  geom_pointrange(aes(x = distmetcat, y = meanIBD, ymin = L95CI, ymax = U95CI)) +
  scale_size(range = c(0, 1)) +
  scale_y_continuous("Mean IBD") +
  facet_wrap(distcat~.) +
  plot_theme +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        axis.text.x = element_text(angle = 90, size = 8, face = "bold"),
        axis.line.x = element_line(size = 1.1, color = "#000000"),
        panel.grid.minor.x = element_line(color = "#bdbdbd"),
        panel.grid.minor.y = element_line(color = "#bdbdbd")
        )


jpeg("results/figures/IBD_decay_plot_meanIBD.jpg",
     unit = "in", height = 8, width = 11, res = 500)
plot(IBDdecay.plot)
graphics.off()

```

```{r, results='asis'}
plot(IBDdecay.plot)
```


## IBD Decay by Province
Note, here we are including all pairwise comparisons for a given sample in a province. 

### IBD Decay Plots - Continuous, Facet Province {.tabset .tabset-fade}

#### All
```{r}

ibDdist.long <- expand_distance_matrix(ibDdist) %>% 
  dplyr::rename(adm1name = adm1name.x)


IBDdecay.df <- ibDdist.long %>%
  dplyr::select(c("smpl1", "smpl2", "malecotf", "adm1name", dplyr::ends_with("cat"))) %>%
  tidyr::gather(., key = "distcat", value = "distmetric", 5:7) %>%
  dplyr::mutate(distcat = factor(distcat, levels = c("gcdistance_cat", "roaddistance_cat",
                                                     "airportdistance_cat"), labels = c("GC Dist.", "Road Dist.", "Airport Dist.")),
                distmetric = factor(distmetric, levels = c("Within", "100km",
                                           "200km", "300km", "400km", "500km", "600km",
                                           "700km", "800km", "900km", "1000km", "1100km",
                                           "1200km", "1300km", "1400km", "1500km", "1600km",
                                           "1700km", "1800km", "1900km", "2000km", ">2000km"
                                           )))

```

```{r}
IBDdecay.plot <- IBDdecay.df %>%
  ggplot() +
  geom_boxplot(aes(x=distmetric, y = malecotf)) +
  facet_grid(adm1name~distcat, scales = "free") +
  plot_theme +
  theme(axis.text.x = element_text(angle = 90))
```

```{r, results='asis', fig.width=11, fig.height=25}
plot(IBDdecay.plot)
```
```{r}
jpeg("results/figures/IBD_decay_plot_continuous_adm1facet.jpg",
     unit = "in", height = 25, width = 11, res = 500)
plot(IBDdecay.plot)
graphics.off()
```

#### GC Distance
```{r}
IBDdecay.plot <- IBDdecay.df %>%
  dplyr::filter(distcat == "GC Dist.") %>% 
  ggplot() +
  geom_boxplot(aes(x=distmetric, y = malecotf)) +
  facet_wrap(adm1name~., scales = "free") +
  plot_theme +
  theme(axis.text.x = element_text(angle = 90))
```

```{r, results='asis', fig.height=11, fig.width=11}
plot(IBDdecay.plot)
```

```{r}
jpeg("results/figures/IBD_decay_plot_continuous_adm1facet_gcdistance.jpg",
     unit = "in", height = 11, width = 8, res = 500)
plot(IBDdecay.plot)
graphics.off()
```

#### Road Distance
```{r}
IBDdecay.plot <- IBDdecay.df %>%
  dplyr::filter(distcat == "Road Dist.") %>% 
  ggplot() +
  geom_boxplot(aes(x=distmetric, y = malecotf)) +
  facet_wrap(adm1name~., scales = "free") +
  plot_theme +
  theme(axis.text.x = element_text(angle = 90))
```

```{r, results='asis', fig.height=11, fig.width=11}
plot(IBDdecay.plot)
```

```{r}
jpeg("results/figures/IBD_decay_plot_continuous_adm1facet_roaddistance.jpg",
     unit = "in", height = 11, width = 8, res = 500)
plot(IBDdecay.plot)
graphics.off()
```

#### Airport Distance
```{r}
IBDdecay.plot <- IBDdecay.df %>%
  dplyr::filter(distcat == "Airport Dist.") %>% 
  ggplot() +
  geom_boxplot(aes(x=distmetric, y = malecotf)) +
  facet_wrap(adm1name~., scales = "free") +
  plot_theme +
  theme(axis.text.x = element_text(angle = 90))

```

```{r, results='asis', fig.height=11, fig.width=11}
plot(IBDdecay.plot)

```

```{r}
jpeg("results/figures/IBD_decay_plot_continuous_adm1facet_airportdistanceance.jpg",
     unit = "in", height = 11, width = 8, res = 500)
plot(IBDdecay.plot)
graphics.off()

```

### 


### IBD Decay Plots - Means, Facet Province {.tabset .tabset-fade}

```{r}

IBDdecay.df <- ibDdist.long %>%
  dplyr::select(c("smpl1", "smpl2", "malecotf", "adm1name", dplyr::ends_with("distance"), "airportdistance")) %>%
  tidyr::gather(., key = "distcat", value = "distmetric", 5:7) %>% 
  dplyr::mutate(distmetcat = cut(x = distmetric, breaks = c(0, 1e-26,  
                                                            seq(100, 2000, by = 100),
                                                            Inf),
                                 right = F, #
                                 labels = c("Within", "100km",
                                            "200km", "300km", "400km", "500km", "600km",
                                            "700km", "800km", "900km", "1000km", "1100km",
                                            "1200km", "1300km", "1400km", "1500km", "1600km",
                                            "1700km", "1800km", "1900km", "2000km", ">2000km")
                                 ),
                distcat = factor(distcat, levels = c("gcdistance", "roaddistance",
                                                    "airportdistance"), 
                                 labels = c("GC Dist.", "Road Dist.", "Airport Dist."))
                ) %>% 
  dplyr::group_by(adm1name, distcat, distmetcat) %>% 
  dplyr::summarise(
    n = n(),
    meanIBD = mean(malecotf),
    sdIBD = sd(malecotf),
    seIBD = sdIBD/sqrt(n),
    U95CI = meanIBD + 1.96 * seIBD,
    L95CI = meanIBD - 1.96 * seIBD
  )
  
  
  
```

#### All
**Let's zoom in a bit since the SE is large for some points**. 
```{r}
  # geom_smooth(aes(x = distmetcat, y = meanIBD), # this needs to be a continuous x -- lift over cat
  #             method="glm",
  #             formula = y+.Machine$double.xmin ~ x,
  #             method.args=list(family=gaussian(link="log")), color = "#9ecae1") +

IBDdecay.plot <- IBDdecay.df %>%
  ggplot() +
  geom_pointrange(aes(x = distmetcat, y = meanIBD, ymin = L95CI, ymax = U95CI)) + 
  scale_size(range = c(0, 1)) +
  scale_y_continuous("Mean IBD") +
  facet_wrap(adm1name ~ distcat) +
  coord_cartesian(ylim = c(0,0.1)) +
  plot_theme +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        axis.text.x = element_text(angle = 90, size = 8, face = "bold"),
        axis.line.x = element_line(size = 1.1, color = "#000000"),
        panel.grid.minor.x = element_line(color = "#bdbdbd"),
        panel.grid.minor.y = element_line(color = "#bdbdbd")
        )


jpeg("results/figures/IBD_decay_plot_meanIBD_adm1.jpg",
     unit = "in", height = 25, width = 11, res = 500)
plot(IBDdecay.plot)
graphics.off()
```

```{r, results='asis', fig.width=11, fig.height=25}
plot(IBDdecay.plot)
```

#### GC Distance
```{r}
  # geom_smooth(aes(x = distmetcat, y = meanIBD), # this needs to be a continuous x -- lift over cat
  #             method="glm",
  #             formula = y+.Machine$double.xmin ~ x,
  #             method.args=list(family=gaussian(link="log")), color = "#9ecae1") +

IBDdecay.plot <- IBDdecay.df %>%
  dplyr::filter(distcat == "GC Dist.") %>% 
  ggplot() +
  geom_pointrange(aes(x = distmetcat, y = meanIBD, ymin = L95CI, ymax = U95CI)) + 
  scale_size(range = c(0, 1)) +
  scale_y_continuous("Mean IBD") +
  facet_wrap(adm1name ~ .) +
  coord_cartesian(ylim = c(0,0.1)) +
  plot_theme +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        axis.text.x = element_text(angle = 90, size = 8, face = "bold"),
        axis.line.x = element_line(size = 1.1, color = "#000000"),
        panel.grid.minor.x = element_line(color = "#bdbdbd"),
        panel.grid.minor.y = element_line(color = "#bdbdbd")
        )


jpeg("results/figures/IBD_decay_plot_meanIBD_adm1_gcdistance.jpg",
     unit = "in", height = 11, width = 8, res = 500)
plot(IBDdecay.plot)
graphics.off()
```

```{r, results='asis', fig.width=11, fig.height=11, fig.align='center'}
plot(IBDdecay.plot)
```


#### Road Distance
```{r}
  # geom_smooth(aes(x = distmetcat, y = meanIBD), # this needs to be a continuous x -- lift over cat
  #             method="glm",
  #             formula = y+.Machine$double.xmin ~ x,
  #             method.args=list(family=gaussian(link="log")), color = "#9ecae1") +

IBDdecay.plot <- IBDdecay.df %>%
  dplyr::filter(distcat == "Road Dist.") %>% 
  ggplot() +
  geom_pointrange(aes(x = distmetcat, y = meanIBD, ymin = L95CI, ymax = U95CI)) + 
  scale_size(range = c(0, 1)) +
  scale_y_continuous("Mean IBD") +
  facet_wrap(adm1name ~ .) +
  coord_cartesian(ylim = c(0,0.1)) +
  plot_theme +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        axis.text.x = element_text(angle = 90, size = 8, face = "bold"),
        axis.line.x = element_line(size = 1.1, color = "#000000"),
        panel.grid.minor.x = element_line(color = "#bdbdbd"),
        panel.grid.minor.y = element_line(color = "#bdbdbd")
        )


jpeg("results/figures/IBD_decay_plot_meanIBD_adm1_roaddistance.jpg",
     unit = "in", height = 11, width = 8, res = 500)
plot(IBDdecay.plot)
graphics.off()
```

```{r, results='asis', fig.width=11, fig.height=11, fig.align='center'}
plot(IBDdecay.plot)
```


#### Airport Distance
```{r}
  # geom_smooth(aes(x = distmetcat, y = meanIBD), # this needs to be a continuous x -- lift over cat
  #             method="glm",
  #             formula = y+.Machine$double.xmin ~ x,
  #             method.args=list(family=gaussian(link="log")), color = "#9ecae1") +

IBDdecay.plot <- IBDdecay.df %>%
  dplyr::filter(distcat == "Airport Dist.") %>% 
  ggplot() +
  geom_pointrange(aes(x = distmetcat, y = meanIBD, ymin = L95CI, ymax = U95CI)) + 
  scale_size(range = c(0, 1)) +
  scale_y_continuous("Mean IBD") +
  facet_wrap(adm1name ~ .) +
  coord_cartesian(ylim = c(0,0.1)) +
  plot_theme +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        axis.text.x = element_text(angle = 90, size = 8, face = "bold"),
        axis.line.x = element_line(size = 1.1, color = "#000000"),
        panel.grid.minor.x = element_line(color = "#bdbdbd"),
        panel.grid.minor.y = element_line(color = "#bdbdbd")
        )


jpeg("results/figures/IBD_decay_plot_meanIBD_adm1_airportdistance.jpg",
     unit = "in", height = 11, width = 8, res = 500)
plot(IBDdecay.plot)
graphics.off()
```

```{r, results='asis', fig.width=11, fig.height=11, fig.align='center'}
plot(IBDdecay.plot)
```

### 

