---
output: html_document
editor_options: 
  chunk_output_type: console
---

# IBD, Meiotic Siblings
In this analysis, we will look at pairs that have more  relatedness than would be expected for a recombining pathogen (e.g. greater than or equal to 0.5 relatedness, so at least meiotic siblings).

```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = T, results = 'hide',
                      fig.align = 'center', fig.width = 8, fig.height = 8)
knitr::opts_knit$set(root.dir = here::here())
```

```{r}
library(tidyverse)
library(raster)
library(tidygraph)
library(ggraph)
library(cowplot)
source("R/themes.R")
source("R/basics.R")
source("R/pairwise_helpers.R")
# load pretty map aesthetics 
DRCprov <- sf::st_as_sf(readRDS("data/map_bases/gadm/gadm36_COD_1_sp.rds"))
load("data/map_bases/space_mips_maps_bases.rda")

# highly related pairs
ibD <- readRDS("data/derived_data/bigbarcode_genetic_data/mipanalyzer.DRCibD.long.mtdt.rds")
ibD.meiotic <- ibD %>%
  dplyr::select(c("smpl1", "smpl2", "hv001.x", "hv001.y", "malecotf")) %>%
  dplyr::filter(malecotf >= 0.5)
# dhs in data
dhsind <- readRDS("data/derived_data/DHS_qPCR_DRCsmpls_geo.rds")


#......................
# get coord points and jitter slightly for plotting
#......................
# load data
mtdt <- readRDS("data/derived_data/sample_metadata.rds") %>% 
  dplyr::select(c("name", "barcode", "hv001", "longnum", "latnum"))

coords.pts <-  mtdt %>% 
  dplyr::filter(hv001 %in% unique(c(ibD.meiotic$hv001.x, ibD.meiotic$hv001.y))) %>% 
  dplyr::mutate(long_jitter = longnum + purrr::map_dbl(dplyr::n(), function(x) rnorm(1, 0.5, 0.01)),
                lat_jitter = latnum + purrr::map_dbl(dplyr::n(), function(x) rnorm(1, 0.5, 01)))


```

## Overview of Meiotic Pairs
### Pair Plot
```{r, results='asis'}
hv001.x <- coords.pts %>% 
  dplyr::rename(hv001.x = hv001) %>% 
  dplyr::select(c("hv001.x", "long_jitter", "lat_jitter"))
hv001.y <- coords.pts %>% 
  dplyr::rename(hv001.y = hv001)  %>% 
  dplyr::select(c("hv001.y", "long_jitter", "lat_jitter"))

ibD.meiotic.hv001 <- dplyr::left_join(ibD.meiotic, hv001.x, by = "hv001.x")
ibD.meiotic.hv001 <- dplyr::left_join(ibD.meiotic.hv001, hv001.y, by = "hv001.y")
ibD.meiotic.hv001 <- ibD.meiotic.hv001 %>% 
  dplyr::mutate(long_jitter.x = ifelse(hv001.x == hv001.y, long_jitter.x - 1e-3, long_jitter.x))
ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  smpl_base_map +
  geom_curve(data = ibD.meiotic.hv001, alpha = 0.5, size = 1.1,
             aes(x = long_jitter.x, y = lat_jitter.x, 
                 xend = long_jitter.y, yend = lat_jitter.y, 
                 color = malecotf)) +
  geom_point(data = coords.pts, aes(x = long_jitter, y = lat_jitter), 
             color = "#d9d9d9", size = 2, show.legend = F, alpha = 0.8) +
  scale_color_viridis_c("Pairwise IBD", option="plasma", direction = 1) 


```

### Pair Table
```{r, results = 'asis'}

ibD.meiotic %>% 
  dplyr::select(c("hv001.x", "smpl1", "hv001.y", "smpl2", "malecotf")) %>% 
  DT::datatable(., extensions='Buttons',
                options = list(
                  searching = T,
                  pageLength = 8,
                  dom = 'Bfrtip', 
                  buttons = c('csv')))

```

### Pairs by Cluster 
```{r, results = 'asis'}
clsts <- mtdt %>% 
  dplyr::select("hv001", "longnum", "latnum") %>% 
  dplyr::filter(!duplicated(.))

ibD.meiotic.clsts <- tibble::tibble(
  hv001 = c(ibD.meiotic$hv001.x, ibD.meiotic$hv001.y)
) %>% 
  dplyr::group_by(hv001) %>% 
  dplyr::summarise(
    ind_within_pairs = n()
  )

ibD.meiotic.clsts <- dplyr::left_join(clsts, ibD.meiotic.clsts, by = "hv001") %>% 
  dplyr::mutate(ind_within_pairs = ifelse(is.na(ind_within_pairs), 0, ind_within_pairs))

ibD.meiotic.clsts <- ibD.meiotic.clsts %>% 
  dplyr::mutate(lvl = ifelse(ind_within_pairs == 0, "N", "Y"))


```

```{r, results='asis'}
ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  smpl_base_map +
  geom_point(data = ibD.meiotic.clsts, aes(x = longnum, y = latnum, 
                                           shape = lvl, fill = ind_within_pairs), 
             alpha = 0.5, color = "#ffffff", size = 1.5) +
  scale_fill_distiller("Count of \n Highly Related Pairs", type = "div", palette = "RdYlBu") +
  scale_shape_manual("Pairs Present", values = c(4, 21)) +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold", size = 12, vjust = 0.5, hjust = 0.5),
    legend.text = element_text(face = "bold", size = 11, angle = 0),
    legend.key = element_rect(fill = "#737373")
  ) + 
  guides(size = guide_legend(ncol=2),
         color = guide_legend(nrow=1)) 

ibD.meiotic.clsts %>% 
  dplyr::filter(lvl == "Y") %>% 
  ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  smpl_base_map +
  geom_point(aes(x = longnum, y = latnum, 
                 shape = lvl, fill = ind_within_pairs), 
             alpha = 0.5, color = "#ffffff", size = 1.5) +
  scale_fill_distiller("Count of \n Highly Related Pairs", type = "div", palette = "RdYlBu") +
  scale_shape_manual("Pairs Present", values = c(21)) +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold", size = 12, vjust = 0.5, hjust = 0.5),
    legend.text = element_text(face = "bold", size = 11, angle = 0),
    legend.key = element_rect(fill = "#737373")
  ) + 
  guides(size = guide_legend(ncol=2),
         color = guide_legend(nrow=1)) 


```
Overall, we identified `r sum(ibD.meiotic.clsts$lvl == "Y")` clusters with at least one sample in a highly related pair while `r sum(ibD.meiotic.clsts$lvl == "N")` clusters had no highly related pairs.

## Connectivity by Geography - Edge Density

```{r}
# convert to network
ibD.meiotic.network <- ibD.meiotic %>% 
  tidygraph::as_tbl_graph(., directed = F) %>% 
  tidygraph::activate("nodes") %>% 
  dplyr::left_join(., mtdt, by = "name")

# edges
meiotic.edges <- ibD.meiotic.network %>% 
  tidygraph::activate("edges") %>% 
  tibble::as_tibble()


# nodes that only have one connection
meiotic_clst.uni <- ibD.meiotic.network %>% 
  tidygraph::activate("nodes") %>%
  dplyr::mutate(node_edge_density = igraph::degree(.)) %>%
  dplyr::filter(c(node_edge_density == 1 | 
                    name %in% c("3002B1W5E", "Z2H3M")# catch violation of transitivity 
  ) 
  )

meiotic_clst.uni.plotobj <- meiotic_clst.uni %>% 
  ggraph(layout = "kk") +
  #geom_node_point(aes(shape = factor(hv001), fill = factor(hv001)), show.legend = F) +
  geom_edge_link(aes(color  = malecotf), width = 1.2, alpha = 0.5) +
  scale_edge_color_viridis("Pairwise IBD") +
  scale_shape_manual(values = rep(c(21, 22, 23, 24, 25), 12)) +
  geom_node_text(aes(label = hv001)) +
  theme_graph() +
  theme(legend.position = "right",
        legend.title = element_text(family = "Helvetica", vjust = 0.85, hjust = 0.5, size = 13, face = "bold"),
        legend.text = element_text(family = "Helvetica", hjust = 1, size = 11, face = "bold"))

# nodes with multiple connections
meiotic_clst.conn <- ibD.meiotic.network %>% 
  tidygraph::activate("nodes") %>% 
  dplyr::mutate(node_edge_density = igraph::degree(.)) %>% 
  dplyr::filter(node_edge_density > 1) 

meiotic_clst.conn.PlotObj <- meiotic_clst.conn %>% 
  ggraph(layout = "kk") +
  #geom_node_point(aes(shape = factor(hv001), fill = factor(hv001)), show.legend = F) +
  geom_edge_link(aes(color  = malecotf), width = 1.2, alpha = 0.5) +
  scale_edge_color_viridis("Pairwise IBD") +
  scale_shape_manual(values = rep(c(21, 22, 23, 24, 25), 12)) +
  geom_node_text(aes(label = hv001)) +
  theme_graph() +
  theme(legend.position = "right",
        legend.title = element_text(family = "Helvetica", vjust = 0.85, hjust = 0.5, size = 13, face = "bold"),
        legend.text = element_text(family = "Helvetica", hjust = 1, size = 11, face = "bold"))

#...................... 
# lay this out for potential figure
#......................
meiotic_sib_networkplot <- cowplot::plot_grid(
  meiotic_clst.uni.plotobj + theme(legend.position = "none"),
  meiotic_clst.conn.PlotObj + theme(legend.position = "none"),
  labels = c("(B)", "(C)"),
  align = "v", ncol = 1)


legend <- cowplot::get_legend(meiotic_clst.uni.plotobj +  
                                theme(legend.position = "bottom",
                                      legend.text = element_text(family = "Helvetica", size = 11, face = "bold", angle = 45)) +
                                guides(color = guide_legend(nrow = 1)))

meiotic_sib_networkplot <- cowplot::plot_grid(meiotic_sib_networkplot, legend, ncol = 1, rel_heights = c(1, 0.1))


#............................................................
# Now pull in and look at Edge Density by geography
#...........................................................
EdgeDens <- ibD.meiotic.network %>% 
  tidygraph::activate(., "nodes") %>% 
  dplyr::mutate(node_edge_density = igraph::degree(.)) %>%
  tibble::as_tibble(.) 

# node labels
nodelabels <- EdgeDens %>% 
  dplyr::select(c("longnum", "latnum", "hv001")) %>% 
  dplyr::filter(!duplicated(.))

# edge plot
edge_dens_plot <- EdgeDens %>% 
  dplyr::mutate(node_edge_density_disc = factor(node_edge_density, # this is a count, so discreteize
                                                levels = 1:max(EdgeDens$node_edge_density))) %>% 
  ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  smpl_base_map +
  geom_point(aes(x = longnum, y = latnum, 
                 fill = node_edge_density_disc), 
             alpha = 0.5, size = 2, shape = 21, stroke = 0.3, color = "#d9d9d9") +
  ggrepel::geom_text_repel(data = nodelabels,
                           aes(x = longnum, y = latnum, label = hv001), 
                           size = 3, fontface = "bold", color = "#bdbdbd") +
  scale_fill_manual("Edge Density", values = rev(RColorBrewer::brewer.pal(max(EdgeDens$node_edge_density), 
                                                                          name = "RdYlBu"))) +
  theme(legend.position = "bottom",
        legend.text = element_text(angle = 0))


```


```{r, results='asis'}
cowplot::plot_grid(edge_dens_plot, meiotic_sib_networkplot, 
                   ncol = 2, rel_widths = c(1, 0.75),
                   labels = c("(A)", ""))

# potential manuscript figure
jpeg("results/figures/meiotic_sib_network_num_geo_connections.jpg", width = 11, height = 8, units = "in", res = 800)
cowplot::plot_grid(edge_dens_plot, meiotic_sib_networkplot, 
                   ncol = 2, rel_widths = c(1, 0.75),
                   labels = c("(A)", ""))
graphics.off()

```
Overall, there are `r nrow(meiotic.edges)` highly-related pairs. Among these highly related pairs, `r sum(meiotic.edges$hv001.x == meiotic.edges$hv001.y)` were from the same cluster. There are `r sum(EdgeDens$node_edge_density == 1)` pairs with only a single connection. 


### Quadrad
Is the quadrad from individuals in the same household?
```{r, results='asis'}
# hv002, households
quadradbarcode <- meiotic_clst.conn %>% 
  tidygraph::activate("nodes") %>% 
  tibble::as_tibble(.) %>% 
  dplyr::filter(hv001 == 284) %>% 
  dplyr::pull(barcode)

# 0. HIV Status (HIV03) ; (personal)
# 1. Biological Sex (HV104)
# 2. Age (HV105)
# 3. Main floor material (categorical: HV213)
# 3. Main wall material (categorical: HV214)
# 3. Main roof material (categorical: HV215)
# 5. Base wealth is hv270)
# 6. Highest year of education completed (continous: HV108) 
# 8. Number of Household Members (continuous: HV009)

dhsind %>% 
  dplyr::filter(barcode %in% tolower(quadradbarcode)) %>% 
  magrittr::set_colnames(tolower(colnames(.))) %>% 
  dplyr::select(c("barcode", "hv001", "hv002", "hv009", "hv104", "hv105", "hv270")) %>% 
  magrittr::set_colnames(c("barcode", "cluster", "household", "num_hs_memb", "sex", "age", "wlthcat"))%>% 
  knitr::kable(.)

```


## "Local" vs. "Long-Range" Transmission
```{r}

#............................................................
# classify connections as local, mixed, or long
#...........................................................
cluster_connect_set <- ibD.meiotic.network %>% 
  tidygraph::activate(., "edges") %>% 
  tibble::as_tibble(.)
cluster_connect_set <- cluster_connect_set %>% 
  dplyr::select(c("hv001.x", "hv001.y", "malecotf")) %>% 
  expand_distance_matrix()
# now categorize
cat_conn <- function(n, local, long) {
  if (n == local) {
    return("local")
  } else if (n == long) {
    return("long")
  } else {
    return("mixed")
  }
}
# get categories
cluster_connect_set <- cluster_connect_set %>% 
  dplyr::group_by(hv001.x) %>% 
  dplyr::summarise(n = dplyr::n(),
                   local = sum(hv001.x == hv001.y),
                   long = sum(hv001.x != hv001.y)) %>% 
  dplyr::mutate(conn_cat = purrr::pmap_chr(list(n, local, long), .f = cat_conn),
                conn_cat = factor(conn_cat, levels = c("local", "mixed", "long"), 
                                  labels = c("Local", "Mixed", "Long"))) %>% 
  dplyr::rename(hv001 = hv001.x)


#............................................................
# drc map
#...........................................................
coords.pts <- dplyr::left_join(coords.pts, cluster_connect_set, by = "hv001")
DRCtransmap <- ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  smpl_base_map +
  geom_curve(data = ibD.meiotic.hv001, alpha = 0.5, size = 1.1,
             aes(x = long_jitter.x, y = lat_jitter.x, 
                 xend = long_jitter.y, yend = lat_jitter.y, 
                 color = malecotf)) +
  geom_point(data = coords.pts, aes(x = long_jitter, y = lat_jitter, shape = conn_cat, fill = conn_cat), 
             color = "#f0f0f0", size = 2, alpha = 0.8) +
  scale_color_viridis_c("Pairwise IBD", option="plasma", direction = 1) +
  scale_shape_manual("Transmission Class", values = c(21, 21, 22)) +
  scale_fill_manual("Transmission Class", values = c("#FEC907", "#B7FF50", "#30ADE5")) +
  theme(legend.key = element_rect(color = "#525252", fill = "#525252"))

# potential manuscript figure
jpeg("results/figures/meiotic_sib_network_tranmission.jpg", width = 8, height = 8, units = "in", res = 800)
plot(DRCtransmap)
graphics.off()
```
```{r, results='asis'}
plot(DRCtransmap)
```


### Distances of Long Transmissions
```{r}
#......................
# Import Distance matrix
#...................... 
distancematrix.cluster <- readRDS("data/distance_data/distancematrix_bycluster.rds") %>% 
  dplyr::mutate(hv001.x = as.numeric(hv001.x),
                hv001.y = as.numeric(hv001.y))

ibDdist <- long_distance_matrix_join(x = ibD.meiotic.hv001, y = distancematrix.cluster,
                                     by = c("hv001.x", "hv001.y")) %>%
  dplyr::mutate(gcdistance = gcdistance/1e3,
                roaddistance = roaddistance/1e3,
                airportdistance = airportdistance/1e3)


# take care of cluster diagnonals
ibDdist$gcdistance[ibDdist$hv001.x == ibDdist$hv001.y] <- 0
ibDdist$roaddistance[ibDdist$hv001.x == ibDdist$hv001.y] <- 0
ibDdist$airportdistance[ibDdist$hv001.x == ibDdist$hv001.y] <- 0

#......................
# subset 
#......................
notsameclst.edges <- ibD.meiotic.network %>% 
  tidygraph::activate(., "edges") %>% 
  tibble::as_tibble(.) %>% 
  dplyr::filter(hv001.x != hv001.y)
ibDdist <- ibDdist %>% 
  dplyr::select(c("smpl1", "smpl2", "hv001.x", "hv001.y", "gcdistance", "roaddistance", "airportdistance"))

```
```{r, results='asis'}
dplyr::left_join(notsameclst.edges,
                 ibDdist, by = c("hv001.x", "hv001.y")) %>% 
  dplyr::select(c("gcdistance", "roaddistance", "airportdistance")) %>% 
  tidyr::gather(., key = "distcat", value = "distance") %>% 
  summ_var(., x = "distance", groupingvar = "distcat") %>% 
  pretty_DT_tab(.)

```


## Meiotic Siblings and Urbanicity
```{r}
#...................... 
# lift over cities for plotting
#......................
drccities <- readr::read_csv("data/map_bases/DRC_city_coordinates.csv") %>% 
  dplyr::filter(population >= 5e4)
# liftover for drc cities
drccities <- drccities %>% 
  dplyr::mutate(pop_fact = cut(population, breaks = c(50e3, 100e3, 250e3, 500e3, 1e6, Inf), right = F),
                pop_fact = factor(pop_fact, labels = c("50-100", "100-250", "250-500", "500-1,000", ">1,000")))


#...................... 
# get dhs points again (no jitter)
#......................
ge <- sf::st_as_sf(readRDS("data/raw_data/dhsdata/datasets/CDGE61FL.rds")) %>% 
  magrittr::set_colnames(tolower(colnames(.))) %>% 
  dplyr::rename(hv001 = dhsclust) %>% 
  dplyr::select(c("hv001", "urban_rura"))
sf::st_geometry(ge) <- NULL

# extract out coordinate information and join in urban rural data
urban.pts <- mtdt %>% 
  dplyr::select(c("hv001", "name", "longnum", "latnum")) %>% 
  dplyr::filter(name %in% unique(c(ibD.meiotic$smpl1, ibD.meiotic$smpl2)))

urban.pts <- urban.pts %>% 
  dplyr::filter(!duplicated(hv001)) %>% 
  dplyr::left_join(., ge, by = "hv001") %>% 
  dplyr::mutate(urban_rura_ext = ifelse(urban_rura == "R", 10000, 2000))

urban.pts <- sf::st_as_sf(urban.pts, coords = c("longnum", "latnum"), crs = 4326)

# keep longnum and lantum though
urban.pts <- urban.pts %>% 
  dplyr::mutate(
    longnum = sf::st_coordinates(geometry)[,1],
    latnum = sf::st_coordinates(geometry)[,2]
  )

#...................... 
# read in urbanicity 
#......................
covarstack <- readRDS("~/Documents/GitHub/Space_the_Final_Miptier/data/derived_data/covar_rasterstack_raw.RDS")
urban <- covarstack[["urban"]]

urbanmean <- rep(NA, nrow(urban.pts)) 
for (i in 1:nrow(urban.pts)) {
  urbanmean[i] <-  raster::extract(x = urban,
                                   y = sf::as_Spatial(urban.pts$geometry[i]),
                                   buffer = urban.pts$urban_rura_ext[i],
                                   fun = mean, na.rm = T
  )
}

# final urban obj
urban.pts <- urban.pts %>% 
  dplyr::mutate(longnum = sf::st_coordinates(geometry)[,1],
                latnum = sf::st_coordinates(geometry)[,2],
                urbanmean = urbanmean)
sf::st_geometry(urban.pts) <- NULL

# get coords
meioticcoords <- ibD.meiotic.hv001 %>% 
  dplyr::select(c("hv001.x", "hv001.y", dplyr::starts_with("long"), dplyr::starts_with("lat"))) %>% 
  dplyr::filter(!duplicated(.))

# make plot
ibD.meiotic_plot_urbanrstr <- ibD.meiotic.network %>% 
  tidygraph::activate(., "edges") %>% 
  dplyr::left_join(., y = meioticcoords) %>% 
  tibble::as_tibble(.) %>% 
  ggplot() +
  ggspatial::layer_spatial(data = urban, aes(fill = stat(band1))) +
  scale_fill_distiller("Urbanicity", type = "div", palette = "RdYlBu", na.value = NA) +
  geom_sf(data = DRCprov, color = "#737373", fill = NA, size = 0.05) +
  smpl_base_map +
  geom_curve(alpha = 0.5, size = 1,
             aes(x = long_jitter.x, y = lat_jitter.x, 
                 xend = long_jitter.y, yend = lat_jitter.y, 
                 color = malecotf)) +
  geom_point(data = urban.pts, aes(x = longnum, y = latnum), 
             size = 2, shape = 21, stroke = 0.3, fill = NA) +
  scale_color_viridis_c("Pairwise IBD", option="plasma", direction = 1) 

```

```{r, results='asis'}
plot(ibD.meiotic_plot_urbanrstr)
```




## Meiotic Siblings and Incidence

```{r, results='asis'}
#.............
# Pf Incidence
#.............
parasiterate <- covarstack[["incidence"]]

#...................... 
# extract incidence data per cluster
#......................
# extract out coordinate information and join in incidence data
incid.pts.meiotic <- mtdt %>% 
  dplyr::select(c("hv001", "name", "longnum", "latnum")) %>% 
  dplyr::filter(name %in% unique(c(ibD.meiotic$smpl1, ibD.meiotic$smpl2)))

incid.pts.meiotic <- incid.pts.meiotic %>% 
  dplyr::left_join(., ge, by = "hv001") %>% 
  dplyr::mutate(urban_rura_ext = ifelse(urban_rura == "R", 10000, 2000))

# cluster 54 has missing when just 2km -- extend to 6km 
incid.pts.meiotic$urban_rura_ext[incid.pts.meiotic$hv001 == 54] <- 6000
# convert to geom
incid.pts.meiotic <- sf::st_as_sf(incid.pts.meiotic, coords = c("longnum", "latnum"), crs = 4326)

incidmean <- rep(NA, nrow(incid.pts.meiotic)) 
for (i in 1:nrow(incid.pts.meiotic)) {
  incidmean[i] <-  raster::extract(x = parasiterate,
                                   y = sf::as_Spatial(incid.pts.meiotic$geometry[i]),
                                   buffer = incid.pts.meiotic$urban_rura_ext[i],
                                   fun = mean, na.rm = T
  )
}

incid.pts.meiotic <- incid.pts.meiotic %>% 
  dplyr::mutate(longnum = sf::st_coordinates(geometry)[,1],
                latnum = sf::st_coordinates(geometry)[,2],
                incidmean = incidmean) 
# subset to final object
incid.pts <- incid.pts.meiotic %>% 
  dplyr::select(c("hv001", "longnum", "latnum", "incidmean")) %>% 
  dplyr::filter(!duplicated(.))
```

```{r}

ibD.meiotic_plot_prev <- ibD.meiotic.network %>% 
  tidygraph::activate(., "edges") %>% 
  tibble::as_tibble(.) %>% 
  dplyr::left_join(., y = meioticcoords) %>% 
  ggplot() +
  ggspatial::layer_spatial(data = parasiterate, aes(fill = stat(band1))) +
  scale_fill_distiller("MAP \n Incidence", type = "div", palette = "RdYlBu", na.value = NA) +
  geom_sf(data = DRCprov, color = "#737373", fill = NA, size = 0.05) +
  smpl_base_map +
  geom_curve(alpha = 0.5, size = 1,
             aes(x = long_jitter.x, y = lat_jitter.x, 
                 xend = long_jitter.y, yend = lat_jitter.y, 
                 color = malecotf)) +
  geom_point(data = incid.pts, aes(x = longnum, y = latnum), 
             size = 2, shape = 21, stroke = 0.3, fill = NA) +
  scale_color_viridis_c("Pairwise IBD", option="plasma", direction = 1) 


```
```{r, results='asis'}
plot(ibD.meiotic_plot_prev)
```


```{r}
# potential mauscript figure
# want to move legend around so rewriting this
# urban
ibD.meiotic_plot_urbanrstr <- ibD.meiotic.network %>% 
  tidygraph::activate(., "edges") %>% 
  tibble::as_tibble(.) %>% 
  dplyr::left_join(., y = meioticcoords) %>% 
  ggplot() +
  ggspatial::layer_spatial(data = urban, aes(fill = stat(band1))) +
  scale_fill_distiller("Urbanicity", type = "div", palette = "RdYlBu", na.value = NA) +
  geom_sf(data = DRCprov, color = "#737373", fill = NA, size = 0.05) +
  smpl_base_map +
  geom_curve(alpha = 0.5, size = 1, show.legend = F,
             aes(x = long_jitter.x, y = lat_jitter.x, 
                 xend = long_jitter.y, yend = lat_jitter.y, 
                 color = malecotf)) +
  geom_point(data = urban.pts, aes(x = longnum, y = latnum), 
             size = 2, shape = 21, stroke = 0.3, fill = NA)  +
  scale_color_viridis_c("Pairwise IBD", option="plasma", direction = 1)

# incidence 
ibD.meiotic_plot_incidence <- ibD.meiotic.network %>% 
  tidygraph::activate(., "edges") %>% 
  dplyr::left_join(., y = meioticcoords) %>% 
  tibble::as_tibble(.) %>% 
  ggplot() +
  ggspatial::layer_spatial(data = parasiterate, aes(fill = stat(band1))) +
  scale_fill_distiller("MAP \n Incidence", type = "div", palette = "RdYlBu", na.value = NA) +
  geom_sf(data = DRCprov, color = "#737373", fill = NA, size = 0.05) +
  smpl_base_map +
  geom_curve(alpha = 0.5, size = 1, show.legend = F,
             aes(x = long_jitter.x, y = lat_jitter.x, 
                 xend = long_jitter.y, yend = lat_jitter.y, 
                 color = malecotf)) +
  geom_point(data = incid.pts, aes(x = longnum, y = latnum), 
             size = 2, shape = 21, stroke = 0.3, fill = NA)  +
  scale_color_viridis_c("Pairwise IBD", option="plasma", direction = 1)




# legend dummy plot obj
legend_dummy <- ibD.meiotic.network %>% 
  tidygraph::activate(., "edges") %>% 
  tibble::as_tibble(.) %>% 
  dplyr::left_join(., y = meioticcoords) %>% 
  ggplot() +
  geom_curve(alpha = 0.5, size = 1,
             aes(x = long_jitter.x, y = lat_jitter.x, 
                 xend = long_jitter.y, yend = lat_jitter.y, 
                 color = malecotf)) +
  scale_color_viridis_c("Pairwise IBD", option="plasma", direction = 1) 

legend <- cowplot::get_legend(legend_dummy +  
                                theme(legend.position = "bottom",
                                      legend.text = element_text(family = "Helvetica", size = 11, face = "bold", angle = 45)))

#...................... 
# bring together
#......................
firstrow <- cowplot::plot_grid(ibD.meiotic_plot_urbanrstr, ibD.meiotic_plot_incidence,
                               nrow = 1, labels = c("(A)", "(B)"), align = "h")
meiotic_sib_covarplot <- cowplot::plot_grid(firstrow, legend, ncol = 1, rel_heights = c(1, 0.1))


# figure out
jpeg("results/figures/meiotic_sib_covars.jpg", width = 11, height = 8, units = "in", res = 800)
plot(meiotic_sib_covarplot)
graphics.off()

svglite::svglite("results/figures/meiotic_sib_covars.jpg", width = 11, height = 8)
plot(meiotic_sib_covarplot)
graphics.off()

```
```{r, results='asis'}
plot(meiotic_sib_covarplot)
```
