---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Traditional Epidemiological Approach
```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = T, results = 'hide',
                      fig.align = 'center', fig.width = 8, fig.height = 8)
```
```{r}
#---------------------------------------------------------------------------
# Purpose of this script is to do basic epi analyses
#----------------------------------------------------------------------------
library(tidyverse)
library(MuMIn)
library(merTools)
library(stargazer)
library(raster)
source("~/Documents/GitHub/Space_the_Final_Miptier/R/basics.R")
source("~/Documents/GitHub/Space_the_Final_Miptier/R/themes.R")
source("~/Documents/GitHub/Space_the_Final_Miptier/R/pairwise_helpers.R")

```

```{r}

drcsmpls <- readRDS("~/Documents/GitHub/Space_the_Final_Miptier/data/distance_data/drcsmpls_foruse.rds") %>%
  magrittr::set_colnames(tolower(colnames(.))) %>%
  dplyr::select(c("id", "hv001")) %>% 
  dplyr::rename(name = id)

mtdt <- readRDS("~/Documents/GitHub/Space_the_Final_Miptier/data/derived_data/sample_metadata.rds") %>% 
  magrittr::set_colnames(tolower(colnames(.))) %>%
  dplyr::rename(name = id) %>% 
  dplyr::select(c("name", "country", "hv001", "adm1name", "longnum", "latnum")) %>% 
  dplyr::filter(name %in% drcsmpls$name)

# drc prov for plot
DRCprov <- sf::st_as_sf(readRDS("~/Documents/GitHub/Space_the_Final_Miptier/data/map_bases/gadm/gadm36_COD_1_sp.rds")) 

# load pretty map aesthetics 
load("~/Documents/GitHub/Space_the_Final_Miptier/data/map_bases/space_mips_maps_bases.rda")
```

```{r}
#....................................................................................
# Import Genetic Data
#....................................................................................
ibD <- readRDS("~/Documents/GitHub/Space_the_Final_Miptier/data/derived_data/bigbarcode_genetic_data/mipanalyzer.DRCibD_polarized_biallelic_processed.long.rds")

# log2 IBD measure
ibD <- ibD %>% 
  dplyr::mutate(malecotf_gens = -log2(malecotf),
                malecotf_gens_inv = 1/malecotf_gens)


ibD.long <- expand_distance_matrix(ibD)
```


## Epi Considerations
Here we will analyze IBD using a traditional epidemiological approach. _Outcome_: pairwise IBD comparisons; _Covariates_: Differences among observations; _Spatial Autocorrelation_: Random Intercepts (Province). Gaussian Process (Cluster). _Link/family_: Given that we can normalize the IBD pairwise outcome, we will use a log-link and the Gaussian probability family for our generalized linear models. A log-link will be used given the expected exponential relationship of IBD and distance/covariates.   
  
NB, all covariates were taken from open sources including: Malaria Atlas Project, European Space Agency, LAADS/NASA, and OSM (this include prevalence which is not from our lab but from MAP).



### Outcome
When considering pairwise IBD as the outcome, we can use a log-base-2 tranformation to normalize the IBD measure. Log-base-2 represents the number of generations ago to samples were IBD (or coalesced) under an assumption of outcrossing. N.B. larger numbers mean more generations (less related).
```{r, results='asis'}
ibD.long %>% 
  dplyr::filter(malecotf_gens != Inf) %>% 
  ggplot() + 
  geom_histogram(aes(x = malecotf_gens, y = ..count.. )) + 
  plot_theme +
  xlab("Generations") + ylab("")

```
Looks relatively normal considering... 


## Province Models
### Covariate Considerations
```{r}
#..............................................................
# Get Covariates for Province
#..............................................................
covarrstr <- readRDS("~/Documents/GitHub/Space_the_Final_Miptier/data/derived_data/covar_rasterstack_raw.RDS")

extract_agg_raster_polygon <- function(rstrlyr, plygn){
  vals <- raster::extract(x = rstrlyr, y = sf::as_Spatial(plygn),
                          fun = mean,
                          na.rm = T, 
                          sp = F
                          )
  return(as.vector(vals))
  
}

adm1 <- DRCprov %>% 
  dplyr::select(c("adm1name", "geometry"))
adm1.list <- split(adm1, 1:nrow(adm1))

provCovar <- lapply(adm1.list, extract_agg_raster_polygon, rstrlyr = covarrstr)
provCovar <- do.call("rbind.data.frame", provCovar)
colnames(provCovar) <- names(covarrstr)

# add back in geometry and ad1names
provCovar <- cbind.data.frame(adm1, provCovar)
provCovar <- sf::st_as_sf(provCovar)

```

Here you can visualize the covariate by province. 
```{r}
# this nonsense since sf and tidyr get upset 
provCovar.nosf <- provCovar
sf::st_geometry(provCovar.nosf) <- NULL

provCovar.geo <- provCovar %>% 
  dplyr::select(c("adm1name", "geometry"))

provCovar.nosf <- provCovar.nosf %>% 
  tidyr::gather(., key = "covar", value = "val", 2:ncol(.)) 
provCovar.geo <- dplyr::left_join(provCovar.nosf, 
                                  y = provCovar.geo, by = "adm1name") 

provCovar.geo <- sf::st_as_sf(provCovar.geo)

# end of nonsense
provCovar.geo %>% 
  ggplot() + 
  geom_sf(aes(fill = val)) +
  facet_wrap(covar ~ .) + 
  scale_fill_distiller("Prevalence", type = "div", palette = "RdYlBu") +
  map_theme +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())

```

### Distribution of IBD-Gens by Prov
```{r, results='asis'}

p1 <- mtdt %>% 
  dplyr::select(c("name", "adm1name")) %>% 
  dplyr::rename(smpl1 = name)

p2 <- mtdt %>% 
  dplyr::select(c("name", "adm1name")) %>% 
  dplyr::rename(smpl2 = name)

ibD.long %>% 
  dplyr::left_join(., y = p1, by = "smpl1") %>% 
  dplyr::left_join(., y = p2, by = "smpl2") %>% 
  expand_distance_matrix() %>% 
  dplyr::filter(malecotf_gens != Inf) %>% 
  ggplot() + 
  ggridges::stat_density_ridges(
    geom = "density_ridges_gradient", calc_ecdf = T,
    quantiles = 4, quantile_lines = T, 
    aes(x = malecotf_gens, y = adm1name.x, fill = factor(..quantile..))
  ) +
  scale_fill_viridis_d("Quantiles") +
  ylab("Province") + xlab("Distribution of Generations") +
  plot_theme 
  
```

## Between Province IBD Models 
Going to fixed mixed models here using a random intercept for Province and a random slope for **geographical space** while also allowing for correlation allowing between the two (in `lmer/lme4` speak: `Y ~ (1+Space|Prov) + Covars`). This will then model the effect of "space" with respect to Province as random effects with the other covariates considered as fixed effects.  The model is formulated as: 

$$ln(IBD_{ij}) = (b_{0j}(s) + \beta_0) + (b_{1j}(s) + \beta^{\intercal})x_{ij} + \epsilon_{ij}$$
Here, $b$ are the random intercepts and slope that rely on space, $s$. Note, $IBD_{ij}$ is calculated at the mean IBD between province $i$ and province $j$. 
  
We will fit these models first (saturated and base for each distance).
```{r}
mods <- readRDS("~/Documents/GitHub/Space_the_Final_Miptier/results/trad_epi_fits/Provglmm_between_modfits.RDS")
```


#####  Model Fits
```{r, results='asis'}
stargazer::stargazer(mods$fitout, 
                     no.space=F,
                     ci=F, ci.level=0.95, 
                     single.row=F,
                     align=T, 
                     type = "html")

```

##### Covar P-values
```{r, results='asis'}
stargazer(mods$fitout,
          title = "Pvalues of Coef from GLMMs",
          report=('vc*p'),
          align=T, 
          type = "html")

```

#####  Model Summarys
(open code chunk below)
```{r}

for(i in 1:nrow(mods)){
  merTools::fastdisp(mods$fitout[[i]])
  cat("\n ******************************************************************* \n")
}



```

##### Between Prov Model Dredging  {.tabset .tabset-fade}

###### Greater Circle Distance Dredge

```{r, results='asis'}
as.data.frame(mods$dredging[mods$class == "full" & mods$space == "gcdistance.scale"]) %>% 
  DT::datatable(.,
                rownames = F,
                options = list(
                  searching = F,
                  pageLength = 6,
                  dom = 'Bfrtip',
                  buttons = c('csv')))



stargazer(mods$dredging[mods$class == "full" & mods$space == "gcdistance.scale"],
          align=T, 
          type = "html")

```

###### Road Distance Dredge

```{r, results='asis'}
as.data.frame(mods$dredging[mods$class == "full" & mods$space == "roaddistance.scale"]) %>% 
  DT::datatable(.,
                rownames = F,
                options = list(
                  searching = F,
                  pageLength = 6,
                  dom = 'Bfrtip',
                  buttons = c('csv')))



stargazer(mods$dredging[mods$class == "full" & mods$space == "roaddistance.scale"],
          align=T, 
          type = "html")

```

###### River Distance Dredge

```{r, results='asis'}
as.data.frame(mods$dredging[mods$class == "full" & mods$space == "riverdistance.scale"]) %>% 
  DT::datatable(.,
                rownames = F,
                options = list(
                  searching = F,
                  pageLength = 6,
                  dom = 'Bfrtip',
                  buttons = c('csv')))



stargazer(mods$dredging[mods$class == "full" & mods$space == "riverdistance.scale"],
          align=T, 
          type = "html")

```


##### 


## Within Province IBD Models 
Note, I excluded weights here. Can add them back in if we decide to use these models.
```{r}
mods <- readRDS("~/Documents/GitHub/Space_the_Final_Miptier/results/trad_epi_fits/Provglm_within_modfits.RDS")
```


#####  Model Fits
```{r, results='asis'}
stargazer::stargazer(mods$fitout, 
                     no.space=F,
                     ci=F, ci.level=0.95, 
                     single.row=F,
                     align=T, 
                     type = "html")

```

##### Covar P-values
```{r, results='asis'}
stargazer(mods$fitout,
          title = "Pvalues of Coef from GLMs",
          report=('vc*p'),
          align=T, 
          type = "html")

```


##### Within Prov Model Dredging  {.tabset .tabset-fade}

###### Within-within Prov Dredge

```{r, results='asis'}
as.data.frame(mods$dredging[mods$class == "covars" & mods$outcome == "wthnIBD"]) %>% 
  DT::datatable(.,
                rownames = F,
                options = list(
                  searching = F,
                  pageLength = 6,
                  dom = 'Bfrtip',
                  buttons = c('csv')))



stargazer(mods$dredging[mods$class == "covars" & mods$outcome == "wthnIBD"],
          align=T, 
          type = "html")

```

###### Mean IBD within Prov Dredge

```{r, results='asis'}
as.data.frame(mods$dredging[mods$class == "covars" & mods$outcome == "meanIBD"]) %>% 
  DT::datatable(.,
                rownames = F,
                options = list(
                  searching = F,
                  pageLength = 6,
                  dom = 'Bfrtip',
                  buttons = c('csv')))



stargazer(mods$dredging[mods$class == "covars" & mods$outcome == "meanIBD"],
          align=T, 
          type = "html")

```


##### 

