---
output: html_document
editor_options: 
  chunk_output_type: console
---

# IBD, Meiotic Siblings Sink-Source
In this analysis, we will look at our meiotic pairs and see if we can determine if directionality based on local IBD. 

```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = T, results = 'hide',
                      fig.align = 'center', fig.width = 8, fig.height = 8)
```
```{r}
library(tidyverse)
library(raster)
library(tidygraph)
library(ggraph)
source("~/Documents/GitHub/Space_the_Final_Miptier/R/themes.R")
```

```{r}
# drc prov for plot
DRCprov <- sf::st_as_sf(readRDS("~/Documents/GitHub/Space_the_Final_Miptier/data/map_bases/gadm/gadm36_COD_1_sp.rds"))
# load pretty map aesthetics 
load("~/Documents/GitHub/Space_the_Final_Miptier/data/map_bases/space_mips_maps_bases.rda")

```
```{r}
mtdt <- readRDS("~/Documents/GitHub/Space_the_Final_Miptier/data/derived_data/sample_metadata.rds")

ibD <- readRDS("~/Documents/GitHub/Space_the_Final_Miptier/data/derived_data/bigbarcode_genetic_data/mipanalyzer.DRCibD.long.mtdt.rds")
# log2 IBD measure
ibD <- ibD %>% 
  dplyr::mutate(malecotf_gens = -log2(malecotf),
                malecotf_gens_inv = 1/malecotf_gens)

```



```{r}
#..............................................................
# Idenity samples with IBD that are greater than half the
# genome (i.e. at or above meiotic siblings)
# These represent likely recent transmission events
#..............................................................
ibD.meiotic <- ibD %>% 
  dplyr::filter(malecotf >= 0.5)
ibdmeioticsmpls <- as.character(ibD.meiotic$smpl1)
ibdmeioticsmpls <- c(ibdmeioticsmpls, as.character(ibD.meiotic$smpl2))


```

## Approach
For each "meiotic" sample, we are going to take the mean IBD within a 10-km radius (about the distance a mosquito flies) of the samples' genotype to the other "near-neighbor" pairs. This should give us a sense of how common that genotype (or really how common segments are) and potentially "root" genotypes. Will scale this by $$\frac{n}{n-1}$$ to account for small sample sizes.

```{r}
# meiotic coord pts
ibdmeioticsmpls.coords.pts <- mtdt %>% 
  dplyr::select(c("name", "hv001", "longnum", "latnum")) %>% 
  dplyr::filter(name %in% ibdmeioticsmpls)
ibdmeioticsmpls.coords.pts <- sf::st_as_sf(ibdmeioticsmpls.coords.pts, 
                                           coords = c("longnum", "latnum"),
                                           crs = sf::st_crs(DRCprov))
ibdmeioticsmpls.coords.pts <- sf::st_transform(ibdmeioticsmpls.coords.pts, 
                                               crs = "+proj=utm +zone=34 +datum=WGS84 +units=m")

#..............................................................
# subset mtdt to just what we need to make a meiotic table
#..............................................................
mtdt.light <- mtdt %>% 
  sf::st_as_sf(., coords = c("longnum", "latnum")) %>% 
  dplyr::select(c("hv001")) %>% 
  dplyr::filter(!duplicated(.))
sf::st_crs(mtdt.light) <- sf::st_crs(DRCprov)
mtdt.light <- sf::st_transform(mtdt.light, crs = "+proj=utm +zone=34 +datum=WGS84 +units=m")

#..............................................................
# find nearest neighbors w/in 10km
# 10-km radius because it corresponds to the maximum flight distance of blood-fed A gambiae
# 10km wasn't enough... so moved on to where we thought we saw an inflection point in the data for falling off of distance approximately 100km 
#..............................................................
meiotic.buff <- sf::st_buffer(ibdmeioticsmpls.coords.pts, 
                              dist = 100*1e3) # m to km
meiotic.buff <- split(meiotic.buff, 1:nrow(meiotic.buff))
ibdmeioticsmpls.coords.pts$clst_intersects <- lapply(meiotic.buff, function(buff){
  ret <- sf::st_intersection(mtdt.light, buff)
  ret <- ret$hv001
  return(ret)
})

# we no longer need geographic information
# as we now have the meiotic sample and which clusters it is closest to 
sf::st_geometry(ibdmeioticsmpls.coords.pts) <- NULL


```

```{r}

#..............................................................
# Making the Meiotic IBD dictionary
#   Here we have the meiotic sibling sample and the clusters
#   that were within X-km of that sample. We extract all pairwise
#   comparison of the meiotic sample and those samples in the nearby clusters
#..............................................................
liftover_ibd_neighborhood <- function(ibd_dict, smpls = drcsmpls, genrelated = ibD){
  # get intersecting clusters
  clst_intersects <- unlist(ibd_dict$clst_intersects)
  # get all drc samples from clsts
  smpls_intersect <- smpls %>% 
    dplyr::filter(hv001 %in% clst_intersects)
  # find pairwise comparison that are for meiotic smpl and neighbors 
  ibD_intersect <- genrelated %>% 
    dplyr::filter( c(
      smpl1 %in% smpls_intersect$name & smpl2 == ibd_dict$name | # transivity of iBD measures
        smpl2 %in% smpls_intersect$name & smpl1 == ibd_dict$name
    )
    )
  # calculate neighborhood IBD
  neighborhood_ibd <- mean(ibD_intersect$malecotf)
  
  return(neighborhood_ibd)
  
}

#..............................................................
# Make this for function
#..............................................................
drcsmpls <- mtdt %>% 
  dplyr::select(c("hv001", "name"))

#..............................................................
# Run Function
#..............................................................
ibdmeioticsmpls.coords.pts.list <- split(ibdmeioticsmpls.coords.pts, 
                                         1:nrow(ibdmeioticsmpls.coords.pts))
ibdmeioticsmpls_neighborIBD <- lapply(ibdmeioticsmpls.coords.pts.list,
                                      liftover_ibd_neighborhood,
                                      smpls = drcsmpls, genrelated = ibD)
```
```{r}
#..............................................................
# Reattach new neighborhood data and join
#..............................................................
ibdmeioticsmpls.coords.pts$neighborhoodIBD = unlist(ibdmeioticsmpls_neighborIBD)
ibdmeioticsmpls.neighborhood <- ibdmeioticsmpls.coords.pts %>% 
  dplyr::select(c("name", "neighborhoodIBD"))

# join with ibd meiotic set
colnames(ibdmeioticsmpls.neighborhood)[1] <- "smpl1"
ibD.meiotic.neighborhoodIBD <- dplyr::left_join(ibD.meiotic, 
                                                ibdmeioticsmpls.neighborhood,
                                                by = "smpl1")
colnames(ibdmeioticsmpls.neighborhood)[1] <- "smpl2"
ibD.meiotic.neighborhoodIBD <- dplyr::left_join(ibD.meiotic.neighborhoodIBD, 
                                                ibdmeioticsmpls.neighborhood,
                                                by = "smpl2")

```

```{r}

ibD.meiotic.neighborhoodIBD <- ibD.meiotic.neighborhoodIBD %>% 
  dplyr::filter(hv001.x != hv001.y) %>% 
  dplyr::mutate(
    neighborIBDdiff = neighborhoodIBD.x - neighborhoodIBD.y # positive values support x, negative y
  )


```

```{r}
# to do, should make this importable since we use it twice
#..............................................................
# Read in urban data
#..............................................................
# extract out coordinate information and join in urban rural data
coord.pts <- sf::st_as_sf(readRDS("~/Documents/GitHub/Space_the_Final_Miptier/data/raw_data/dhsdata/datasets/CDGE61FL.rds")) %>% 
  magrittr::set_colnames(tolower(colnames(.))) %>% 
  dplyr::rename(hv001 = dhsclust) %>% 
  dplyr::select(c("hv001", "urban_rura")) %>% 
  dplyr::filter(!duplicated(hv001)) %>% 
  dplyr::filter(hv001 %in% unique(mtdt$hv001)) %>% 
  dplyr::mutate(urban_rura_ext = ifelse(urban_rura == "R", 10000, 2000))

urban <- raster::raster("~/Documents/GitHub/Space_the_Final_Miptier/data/derived_data/urbanicity_raster/urbanicity.grd")


urbanmedian <- rep(NA, nrow(coord.pts)) 
for (i in 1:nrow(coord.pts)) {
  urbanmedian[i] <-  raster::extract(x = urban,
                                     y = sf::as_Spatial(coord.pts$geometry[i]),
                                     buffer = coord.pts$urban_rura_ext[i],
                                     fun = median, na.rm = T
  )
}

urban.pts <- coord.pts %>% 
  dplyr::mutate(longnum = sf::st_coordinates(geometry)[,1],
                latnum = sf::st_coordinates(geometry)[,2],
                urbanmedian = urbanmedian) %>% 
  dplyr::select(c("hv001", "longnum", "latnum", "urbanmedian")) %>% 
  dplyr::filter(hv001 %in% c(ibD.meiotic.neighborhoodIBD$hv001.x, ibD.meiotic.neighborhoodIBD$hv001.y))


```


```{r}
#..............................................................
# attach urban data
#..............................................................
colnames(urban.pts)[1] <- "hv001.x"
ibD.meiotic.neighborhoodIBD <- dplyr::left_join(ibD.meiotic.neighborhoodIBD,
                                                urban.pts, 
                                                by = "hv001.x")

colnames(urban.pts)[1] <- "hv001.y"
ibD.meiotic.neighborhoodIBD <- dplyr::left_join(ibD.meiotic.neighborhoodIBD,
                                                urban.pts, 
                                                by = "hv001.y")

```

## Directionality Mapping
Here, we have assigned the directionality based on the difference in pairwise IBD between the meiotic smaple and the other samples among local clusters. Arrows indicate the inferred directionality based on local IBD-ness (i.e. if a pairwise comparison shared more similarity in its "from" node versus its "to" node). **N.B. -- subsetted to observations that are between clusters, as we assume that within cluster highly related pairs are likely local transmission events**. 

### Pairwise
```{r}

ibD.meiotic.neighborhoodIBD_plotdf <- ibD.meiotic.neighborhoodIBD %>% 
  # note, need to determine "correct end" for arrow
  dplyr::mutate(long.start = ifelse(neighborIBDdiff > 0, longnum.x, longnum.y),
                lat.start = ifelse(neighborIBDdiff > 0, latnum.x, latnum.y),
                long.end = ifelse(neighborIBDdiff > 0, longnum.y, longnum.x),
                lat.end = ifelse(neighborIBDdiff > 0, latnum.y, latnum.x),
                from = ifelse(neighborIBDdiff > 0, hv001.x, hv001.y),
                to = ifelse(neighborIBDdiff > 0, hv001.y, hv001.x),
                neighborIBDdiff_scaled = abs(neighborIBDdiff),
                neighborIBDdiff_lvls = cut(neighborIBDdiff_scaled,
                                           breaks = c(0, 0.01, 0.025, 0.05, 0.09, 1)),
                neighborIBDdiff_lvls = factor(neighborIBDdiff_lvls, 
                                              labels = c("<1%", "1-2.5%", "2.5-5%", "5-9%", ">9%"))
  )

ibD.meiotic.neighborhoodIBD_pointdf <- mtdt %>% 
  dplyr::filter(hv001 %in% c(ibD.meiotic.neighborhoodIBD_plotdf$hv001.x, ibD.meiotic.neighborhoodIBD_plotdf$hv001.y)) %>% 
  dplyr::select(c("hv001", "longnum", "latnum")) %>% 
  dplyr::filter(!duplicated(.))

IBD_meiotic_direct_plot <- ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  prettybasemap_nodrc_dark +
  geom_curve(data = ibD.meiotic.neighborhoodIBD_plotdf,
             aes(x = long.start, y = lat.start, 
                 xend = long.end, yend = lat.end, 
                 color = malecotf,
                 size = neighborIBDdiff_lvls),
             arrow = arrow(length = unit(0.02, "npc")),
             alpha = 0.5) +
  geom_point(data = ibD.meiotic.neighborhoodIBD_pointdf,
             aes(x = longnum, y = latnum), 
             show.legend = F, shape = 21,
             fill = "#ffffff", color = "#000000" ) + 
  scale_color_viridis_c("Pairwise IBD", option="plasma", direction = 1) +
  scale_size_manual("IBD Nieghborhood \n Differences", values = c(0.4, 0.6, 0.8, 1, 1.2)) +
  map_theme +
  theme(
    legend.position = "right",
    legend.title = element_text(family = "Helvetica", face = "bold", hjust = 0.5, vjust = 0.5, size = 12),
    legend.text = element_text(family = "Helvetica", size = 11, hjust = 0.5, vjust = 0.5, angle = 0),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )  

jpeg("~/Documents/GitHub/Space_the_Final_Miptier/results/figures/DRC_meiotic_sibs_directionality.jpg", width = 8, height = 8, units = "in", res = 500)
plot(IBD_meiotic_direct_plot)
graphics.off()
```

```{r, results='asis'}
plot(IBD_meiotic_direct_plot)
```


### Pairwise-Urbanicity
```{r}


IBD_meiotic_direct_plot.urb <- ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  prettybasemap_nodrc_dark +
  geom_curve(data = ibD.meiotic.neighborhoodIBD_plotdf,
             aes(x = long.start, y = lat.start, 
                 xend = long.end, yend = lat.end, 
                 color = malecotf,
                 size = neighborIBDdiff_lvls),
             arrow = arrow(length = unit(0.02, "npc")),
             alpha = 0.5) +
  geom_point(data = urban.pts,
             aes(x = longnum, y = latnum,
                 fill = urbanmedian), 
             shape = 21) +
  scale_color_viridis_c("Pairwise IBD", option="plasma", direction = -1) +
  scale_size_manual("IBD Nieghborhood \n Differences", values = c(0.4, 0.6, 0.8, 1, 1.2)) +
  scale_fill_distiller("Urbanicity", type = "div", palette = "RdYlBu") +
  map_theme +
  theme(
    legend.position = "right",
    legend.title = element_text(family = "Helvetica", face = "bold", hjust = 0.5, vjust = 0.5, size = 12),
    legend.text = element_text(family = "Helvetica", size = 11, hjust = 0.5, vjust = 0.5, angle = 0),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) 

jpeg("~/Documents/GitHub/Space_the_Final_Miptier/results/figures/DRC_meiotic_sibs_directionality_urban.jpg", width = 8, height = 8, units = "in", res = 500)
plot(IBD_meiotic_direct_plot.urb)
graphics.off()
```

```{r, results='asis'}
plot(IBD_meiotic_direct_plot.urb)
```


### Pairwise Prevalence
```{r}

#..............................................................
# Pf Prevalence
#..............................................................
parasiterate <- raster::raster("~/Documents/GitHub/Space_the_Final_Miptier/data/derived_data/MAPrasters/parasiterate.grd")

# extract out coordinate information and join in prevdata
prev.pts <- sf::st_as_sf(coord.pts, coords = c("longnum", "latnum"), crs = 4326)
# cluster 54 has missing when just 2km -- extend to 6km 
prev.pts$urban_rura_ext[prev.pts$hv001 == 54] <- 6000

prevmean <- rep(NA, nrow(prev.pts)) 
for (i in 1:nrow(prev.pts)) {
  prevmean[i] <-  raster::extract(x = parasiterate,
                                  y = sf::as_Spatial(prev.pts$geometry[i]),
                                  buffer = prev.pts$urban_rura_ext[i],
                                  fun = mean, na.rm = T
  )
}

prev.pts <- prev.pts %>% 
  dplyr::mutate(long = sf::st_coordinates(geometry)[,1],
                lat = sf::st_coordinates(geometry)[,2],
                prevmean = prevmean) %>% 
  dplyr::select(c("hv001", "long", "lat", "prevmean")) %>% 
  dplyr::filter(hv001 %in% mtdt$hv001[mtdt$name %in% ibdmeioticsmpls])

#..............................................................
# Combine data with directionality
#..............................................................
colnames(prev.pts)[1] <- "hv001.x"
ibD.meiotic.neighborhoodIBD_plotdf <- dplyr::left_join(ibD.meiotic.neighborhoodIBD_plotdf, 
                                                       prev.pts, 
                                                       by = "hv001.x")
colnames(prev.pts)[1] <- "hv001.y"
ibD.meiotic.neighborhoodIBD_plotdf <- dplyr::left_join(ibD.meiotic.neighborhoodIBD_plotdf, 
                                                       prev.pts, 
                                                       by = "hv001.y")

```

```{r}
IBD_meiotic_direct_plot.prev <- ggplot() +
  geom_sf(data = DRCprov, color = "#737373", fill = "#525252", size = 0.05) +
  prettybasemap_nodrc_dark +
  geom_curve(data = ibD.meiotic.neighborhoodIBD_plotdf,
             aes(x = long.start, y = lat.start, 
                 xend = long.end, yend = lat.end, 
                 color = malecotf,
                 size = neighborIBDdiff_lvls),
             arrow = arrow(length = unit(0.02, "npc")),
             alpha = 0.5) +
  geom_point(data = prev.pts,
             aes(x = long, y = lat,
                 fill = prevmean), 
             shape = 21) +
  scale_color_viridis_c("Pairwise IBD", option="plasma", direction = 1) +
  scale_size_manual("IBD Nieghborhood \n Differences", values = c(0.4, 0.6, 0.8, 1, 1.2)) +
  scale_fill_distiller("Parasite Rate", type = "div", palette = "RdYlBu") +
  map_theme +
  theme(
    legend.position = "right",
    legend.title = element_text(family = "Helvetica", face = "bold", hjust = 0.5, vjust = 0.5, size = 12),
    legend.text = element_text(family = "Helvetica", size = 11, hjust = 0.5, vjust = 0.5, angle = 0),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) 

jpeg("~/Documents/GitHub/Space_the_Final_Miptier/results/figures/DRC_meiotic_sibs_directionality_prev.jpg", width = 8, height = 8, units = "in", res = 500)
plot(IBD_meiotic_direct_plot.prev)
graphics.off()
```

```{r, results='asis'}
plot(IBD_meiotic_direct_plot.prev)
```


```{r}
jpeg("~/Documents/GitHub/Space_the_Final_Miptier/results/figures/DRC_meiotic_sibs_directionality_facet.jpg", width = 11, height = 6, units = "in", res = 500)
cowplot::plot_grid(IBD_meiotic_direct_plot.urb, IBD_meiotic_direct_plot.prev,
                   labels = c("(A)", "(B)"), align = "h")
graphics.off()
```


### Neighboorhood Data Summary
```{r, results='asis'}
ibD.meiotic.neighborhoodIBD_plotdf %>% 
  dplyr::select(-c(dplyr::starts_with("long"),
                   dplyr::starts_with("lat"),
                   dplyr::starts_with("geometry"),
                   malecotf_gens,
                   malecotf_gens_inv,
                   neighborIBDdiff_scaled,
                   neighborIBDdiff_lvls
  )) %>% 
  dplyr::select(c("hv001.x", "hv001.y", "from", "to", dplyr::everything())) %>% 
  dplyr::rename(
    InferredFrom = from,
    InferredTo = to
  ) %>% 
  dplyr::mutate_if(is.numeric, round, 2) %>% 
  DT::datatable(., extensions='Buttons',
                options = list(
                  searching = T,
                  pageLength = 20,
                  dom = 'Bfrtip', 
                  buttons = c('csv'))) 

```

#### Covariates Based on Directionality

```{r}
# inferred from and to 

tofromcovar <- ibD.meiotic.neighborhoodIBD_plotdf %>% 
  dplyr::mutate(prevfrom = ifelse(hv001.x == from, prevmean.x, prevmean.y),
                prevto = ifelse(hv001.x == to, prevmean.x, prevmean.y),
                urbanfrom = ifelse(hv001.x == from, urbanmedian.x, urbanmedian.y),
                urbanto = ifelse(hv001.x == to, urbanmedian.x, urbanmedian.y)) %>% 
  dplyr::select(c("from", "to", "hv001.x", "hv001.y", "smpl1", "smpl2", "malecotf", "neighborIBDdiff", "prevfrom", "prevto", "urbanfrom", "urbanto")) 

```

###### Prevalence 
```{r, results='asis'}
# histogram of differences
prevdiffhist <- tofromcovar %>% 
  dplyr::mutate(prevdiff = prevfrom - prevto) %>% 
  ggplot() + 
  geom_density(aes(x = prevdiff), fill = "#bdbdbd", color = NA) + 
  geom_vline(xintercept = 0, color = "red") +
  xlab("Prev Diff") + ggtitle("Clst Prev (from) - Clst Prev (to)") + 
  theme_classic()

# from clusters
fromclstprev <- tofromcovar %>% 
  dplyr::select(c("from", "prevfrom")) %>% 
  dplyr::group_by(from) %>% 
  dplyr::summarise(
    edgedens = n(),
    prev = unique(prevfrom) # same cluster by cluster
  )

fromclstprevplotObj <- fromclstprev %>% 
  ggplot() + 
  geom_point(aes(x = edgedens, y = prev, color = factor(from)), show.legend = F) +
  ggtitle("Prev vs. Edge Density From Clst") + 
  theme_classic()

# to clusters
toclstprev <- tofromcovar %>% 
  dplyr::select(c("to", "prevto")) %>% 
  dplyr::group_by(to) %>% 
  dplyr::summarise(
    edgedens = n(),
    prev = unique(prevto) # same cluster by cluster
  )

tolstprevplotObj <- toclstprev %>% 
  ggplot() + 
  geom_point(aes(x = edgedens, y = prev, color = factor(to)), show.legend = F) +
  ggtitle("Prev vs. Edge Density Into Clst") + 
  theme_classic()


cowplot::plot_grid(prevdiffhist, fromclstprevplotObj, tolstprevplotObj,
                   nrow = 1, align = "h")

```


###### Urbanicity 
```{r, results='asis'}
urbandiffhist <- tofromcovar %>% 
  dplyr::mutate(urbandiff = urbanfrom - urbanto) %>% 
  ggplot() + 
  geom_density(aes(x = urbandiff), fill = "#bdbdbd", color = NA) + 
  geom_vline(xintercept = 0, color = "red") +
  xlab("Urban Diff") + ggtitle("Clst Urban (from) - Clst Urban (to)") + 
  theme_classic()

# from clusters
fromclsturban <- tofromcovar %>% 
  dplyr::select(c("from", "urbanfrom")) %>% 
  dplyr::group_by(from) %>% 
  dplyr::summarise(
    edgedens = n(),
    urbanicity = unique(urbanfrom) # same cluster by cluster
  )

fromclsturbanplotObj <- fromclsturban %>% 
  ggplot() + 
  geom_point(aes(x = edgedens, y = urbanicity, color = factor(from)), show.legend = F) +
  ggtitle("Urban vs. Edge Density from Clst") + 
  theme_classic()

# to clusters
toclsturban <- tofromcovar %>% 
  dplyr::select(c("to", "urbanto")) %>% 
  dplyr::group_by(to) %>% 
  dplyr::summarise(
    edgedens = n(),
    urbanicity = unique(urbanto) # same cluster by cluster
  )

tolsturbanplotObj <- toclsturban %>% 
  ggplot() + 
  geom_point(aes(x = edgedens, y = urbanicity, color = factor(to)), show.legend = F) +
  ggtitle("urban vs. Edge Density into Clst") + 
  theme_classic()


cowplot::plot_grid(urbandiffhist, fromclsturbanplotObj, tolsturbanplotObj,
                   nrow = 1, align = "h")


```




###### Urban-Prevalence 

```{r, results='asis'}
library(plotly)
#..............................................................
# From 
#..............................................................
fromcovars <- tofromcovar %>% 
  dplyr::select(c("from", "prevfrom", "urbanfrom")) %>% 
  dplyr::group_by(from) %>% 
  dplyr::summarise(
    edgedens = n(),
    urban = unique(urbanfrom), 
    prev = unique(prevfrom)
  )

from3dplotObj <- plotly::plot_ly(fromcovars, 
                                 x = ~edgedens, 
                                 y = ~urban, 
                                 z = ~prev,
                                 color = ~factor(from)) %>%
  add_markers() %>%
  layout(title = "Nodes are From/Source",
         scene = list(xaxis = list(title = 'Edge Dens'),
                      yaxis = list(title = 'Urban'),
                      zaxis = list(title = 'Prev')))

from3dplotObj
```

```{r, results='asis'}

#..............................................................
# To 
#..............................................................
tocovars <- tofromcovar %>% 
  dplyr::select(c("to", "prevto", "urbanto")) %>% 
  dplyr::group_by(to) %>% 
  dplyr::summarise(
    edgedens = n(),
    urban = unique(urbanto), 
    prev = unique(prevto)
  )

to3dplotObj <- plotly::plot_ly(tocovars, 
                               x = ~edgedens, 
                               y = ~urban, 
                               z = ~prev,
                               color = ~factor(to)) %>%
  add_markers() %>%
  layout(title = "Nodes are To/Sink",
         scene = list(xaxis = list(title = 'Edge Dens'),
                      yaxis = list(title = 'Urban'),
                      zaxis = list(title = 'Prev')))

to3dplotObj


```

###### Neighboorhood Data 

```{r, results='asis'}
tofromcovar %>% 
  dplyr::rename(
    InferredFrom = from,
    InferredTo = to
  ) %>% 
  dplyr::mutate_if(is.numeric, round, 2) %>% 
  DT::datatable(., extensions='Buttons',
                options = list(
                  searching = T,
                  pageLength = 20,
                  dom = 'Bfrtip', 
                  buttons = c('csv'))) 

```


